### 2018-01-20 article/prework.tex
(+) prototype of article

\documentclass{article}
\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{cite}
\usepackage{hyperref}

\title{Airspace configuration with Monte Carlo tree search}
\author{Us}
\date{\today}

\begin{document}
\maketitle

\bibliography{article}
\bibliographystyle{plain}
\end{document}

### 2018-01-23 article/prework.tex
(+) added ref and written content

@@
 \usepackage{hyperref}
+\usepackage{unicode-math}
+\usepackage{amsmath,amsthm,stmaryrd}
 
@@
 
+\section{Preliminaries}
+\subsection{Base theory: the environment and notations}
+Notations based on Markov Decision Problem's theory (taken
+from~\cite{browne2012survey})
+\begin{itemize}
+  \item \(S\) a set of states,
+  \item \(\mathcal{A}\) a set of actions,
+  \item \(T(s, a, s')\) a transition model that determines the probability of
+      reaching state \(s'\) if action \(a\) is applied to state \(s\),
+  \item \(R(s)\) a reward function
+\end{itemize}
+
+\subsection{Monte carlo methods~\cite{browne2012survey}}
+We denote by \(Q(s, a)\) the expected reward of an action. Let
+\begin{itemize}
+  \item \(N(s, a)\) be
+    the number of times action \(a\) has been selected from state \(s\),
+  \item \(N(s)\) the number of times the game has been played from state \(s\),
+  \item \(\mathbb{I}_i(s, a)\) is 1 if action \(a\) was selected from state
+    \(s\) on the \(i\)th playout from state \(s\) else 0 and
+  \item \(z_i\) is the result of the \(i\)th simulation played.
+\end{itemize}
+In Monte Carlo methods, we have
+\begin{equation}
+  Q(s, a) = \frac{1}{N(s, a)}\sum_{i = 1}^{N(s)}\mathbb{I}_i(s, a)z_i
+\end{equation}
+
+\subsection{Monte Carlo tree search}
+In the MCTS, the action \(a\) is the path that leads from the root the the best
+node computed so far
+
 \bibliography{article}

### 2018-01-24 article/prework.tex
(+) added algorithms

@@
 \usepackage{amsmath,amsthm,stmaryrd}
+\usepackage{algorithm,algpseudocode}
 
@@
 \begin{itemize}
-  \item \(S\) a set of states,
+  \item \(\mathcal{S}\) a set of states,
   \item \(\mathcal{A}\) a set of actions,
@@
 \end{itemize}
+
+A policy is a mapping \(\pi \colon \mathcal{S} \to \mathcal{A}\). The aim is to
+output, from a state \(s\), the action that gives the highest reward.
 
@@
 
+\subsection{Bandits methods}
+
 \subsection{Monte Carlo tree search}
-In the MCTS, the action \(a\) is the path that leads from the root the the best
-node computed so far
+In the MCTS, the action \(a\) is the path that leads from the root to the best
+node computed so far. To each node should correspond a state \(s\). This way, an
+action \(a\) from a node with state \(s\) gives an other node with an other
+state \(s'\).
 
@@
 \bibliographystyle{plain}
+
+\section{Algorithms}
+\subsection{Monte carlo}
+\begin{algorithm}
+  \caption{General MCTS approach~\cite{browne2012survey}}
+  \begin{algorithmic}
+    \Function{MctsSearch}{$s_0$}
+    \State{}create root node \(v_0\) with state \(s_0\)
+    \While{within computational budget}
+    \State{} \(v_l \gets\) \Call{TreePolicy}{$v_0$}
+    \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$s(v_l)$}
+    \State{} \Call{Backup}{$v_l$, $\Delta$}
+    \EndWhile{}
+    \Return{\(a(\)\Call{BestChild}{$v_0$}\()\)}
+    \EndFunction{}
+  \end{algorithmic}
+\end{algorithm}
 \end{document}

### 2018-01-25 article/prework.tex
(+) a word on an article

@@
 
-\subsection{Bandits methods}
+\subsection{Bandits methods~\cite{browne2012survey},~\cite{kocsis2006bandit}}
 
@@
 state \(s'\).
-
-\bibliography{article}
-\bibliographystyle{plain}
 
@@
 \end{algorithm}
+
+\section{Per article review}
+\paragraph{Branch and price~\cite{treimuth2016branch}}
+The most important in this article is the model used. The rest depends heavily
+on the algorithm used, which is the branch and price, a variant of the branch
+and bound.
+
+Three things to minimise:
+\begin{itemize}
+  \item atc workload \(a\),
+  \item workload difference between atc \(b\),
+  \item changes between consecutive configurations \(c\).
+\end{itemize}
+The objective function is therefore a linear combination \(\alpha a + \beta b +
+\gamma c\).
+
+The airspace is modelled by a weighted graph (with \(T\) the discretised time)
+\(G = (V, E, W_t, t\in T)\), with
+\begin{itemize}
+  \item \(V\), the set of vertex, elementary sectors of the airspace weighted by
+    its overload,
+  \item \(E\), the set of edges, they connect two contiguous elementary sectors,
+    they are weighted by the coordination workload for the planes crossing the
+    border if the two sectors belong to two different controllers,
+  \item \(W_t\) the weights at time \(t\in T\).
+\end{itemize}
+
+Three sets of binary decisions are introduced, controlling
+\begin{itemize}
+  \item for each feasible configuration, whether it is applied,
+  \item for each edge, whether it is a frontier,
+  \item for each edge, whether its frontier status has changed (shifts from
+    frontier to not frontier and vice versa).
+\end{itemize}
+In each set, there is, per time period, one variable per feasible configuration
+for the first one and one variable per edge for the second and third ones.
+
+The way to generate feasible configurations isn't explained though\dots
+
+\bibliography{article}
+\bibliographystyle{plain}
 \end{document}

### 2018-01-28 article/prework.tex
(+) added tree policy
(+) new article

@@
 
+\subsubsection{Tree policy}
+The tree policy includes the selection and the expansion.
+\begin{algorithm}
+  \caption{Tree policy}
+  \begin{algorithmic}
+    \Function{TreePolicy}{$v$}
+    \While{$v$ is nonterminal}
+    \If{$v$ is expandable}
+    \Return{\Call{Expand}{$v$}}
+    \Else{}{$v \gets$ \Call{BestChild}{$v, C_p$}}
+    \EndIf{}
+    \EndWhile{}
+    \Return{$v$}
+    \EndFunction{}
+  \end{algorithmic}
+\end{algorithm}
+
 \section{Per article review}
@@
 
+\paragraph{Dynamic airspace configuration by genetic
+  algorithm\cite{sergeeva2017dynamic}}
+The cost function uses five criteria,
+\begin{itemize}
+  \item imbalance between the workload of the controlled sectors,
+  \item the coordination workload,
+  \item the number of flight re-entry events,
+  \item the number of short transits flight through sectors,
+  \item the number of controlled sectors in each airpsace configuration.
+\end{itemize}
+Only the first one is relevant here, as the neural network gives the workload of
+each module.
+
+The airspace is modelled as a graph, the problem boils down to a graph
+partitioning problem.
+
+The imbalance is calculated as follow, with \(K_i\) the number of controlled
+sectors at time \(t_i\), \(W_{c_{jk}}\) the total workload of all airspace
+blocks composing the sector \(k\) at time \(t_i\) and \(c\) the targeted
+workload of the sector,
+\[
+  U_t = \sqrt{%
+    \frac{%
+      \sum_{k=1}^{K_i} {\left( \frac{|W_{c_{jk}} - c|}{c} \right)}^2
+    }{k_i}
+  }
+\]
+
 \bibliography{article}

### 2018-01-29 article/article.tex
(+) added article
(+) structure article

--- 2018-01-20 article/prework.tex
@@
-\documentclass{article}
+\documentclass[twoside,twocolumn]{article}
+
 \usepackage{fontspec}
@@
 
-\title{Airspace configuration with Monte Carlo tree search}
-\author{Us}
+\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
+\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
+\usepackage{booktabs}
+
+\usepackage{lettrine}
+
+\usepackage{enumitem}
+\setlist[itemize]{noitemsep}
+
+\usepackage{abstract}
+\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
+\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
+
+\usepackage{titlesec} % Allows customization of titles
+\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
+\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
+\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
+\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
+
+\usepackage{fancyhdr} % Headers and footers
+\pagestyle{fancy} % All pages have headers and footers
+\fancyhead{} % Blank out the default header
+\fancyfoot{} % Blank out the default footer
+\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
+\fancyfoot[RO,LE]{\thepage} % Custom footer text
+
+\usepackage{titling} % Customizing the title section
+
+\usepackage{hyperref} % For hyperlinks in the PDF
+
+\title{Airspace configuration decision based on Monte-Carlo tree search}
+\author{Gabriel~Hondet, Beno\^{\i}t Viry}
 \date{\today}
 
+%----------------------------------------------------------------------------------------
+%	TITLE SECTION
+%----------------------------------------------------------------------------------------
+
+\setlength{\droptitle}{-4\baselineskip} % Move the title up
+
+\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
+\posttitle{\end{center}} % Article title closing formatting
+
+\renewcommand{\maketitlehookd}{%
+\begin{abstract}
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\end{abstract}
+}
+
+%----------------------------------------------------------------------------------------
+
 \begin{document}
+
 \maketitle
+
+\section{Introduction}
 

### 2018-01-30 article/article.tex
(+) began writing article

@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 \end{abstract}
@@
 \section{Introduction}
+The airspace is divided into sectors, themselves divided into elementary modules.
+Currently, configuration is mainly decided on the fly by chief officer (TODO).
+This decision is based on the actual workload on each position. This intuitive
+approach does not take into account future traffic. This papers aims at
+providing a method offering a smoother workload during the day.
+
+Several methods have been considered to solve the dynamic airspace configuration
+problem, for instance via genetic algorithms in~\cite{sergeeve2017dynamic},
+constraint local search in~\cite{jagare2013constraint}, integer linear
+programming in~\cite{treimuth2016branch} or dynamic programming
+in~\cite{bloem2010dynamic}.
+
+To achieve that, as seen in~\cite{treimuth2016branch}
+or~\cite{sergeeva2017dynamic}, a temporal sequence of configurations will be
+provided. Two costs will be considered to create the sequence, namely the cost
+associated to each configuration and transition. The former is based on the
+workload estimated for one sector given a neural network
+(see~\cite{gianazza2010forecasting}).
 

### 2018-01-30 article/article.tex
(+) almost finished intro

@@
 \section{Introduction}
-The airspace is divided into sectors, themselves divided into elementary modules.
+The airspace is divided into sectors, themselves divided into elementary
+modules. Each sector is managed by a controller working position composed of two
+controllers.
+During the day, sectors are split and merged to be able to
+manage the varying traffic. Splitting creates smaller sectors and is therefore
+used when traffic gets too dense. On the opposite, merging sectors allows fewer
+controllers to manage the same airspace, and is therefore used when traffic
+becomes sparse.
+
 Currently, configuration is mainly decided on the fly by chief officer (TODO).
@@
 Several methods have been considered to solve the dynamic airspace configuration
-problem, for instance via genetic algorithms in~\cite{sergeeve2017dynamic},
+problem, for instance via genetic algorithms in~\cite{sergeeva2017dynamic},
 constraint local search in~\cite{jagare2013constraint}, integer linear
@@
 
-To achieve that, as seen in~\cite{treimuth2016branch}
+As seen in~\cite{treimuth2016branch}
 or~\cite{sergeeva2017dynamic}, a temporal sequence of configurations will be
@@
 
+The sequence of configurations will be represented as a tree. The resulting
+sequence will therefore be a path from the root to a leaf of this tree. This
+problem is highly combinatorial (partitioning of the airspace). Knowing that a
+sochastic tree search algorithm has been honing from a good Go
+player~\cite{gelly2012go} to the best one with AlphaGo.
+
 \bibliography{article}

### 2018-01-30 article/article.tex
(+) intro finished

@@
 problem is highly combinatorial (partitioning of the airspace). Knowing that a
-sochastic tree search algorithm has been honing from a good Go
-player~\cite{gelly2012go} to the best one with AlphaGo.
+sochastic tree search algorithm has been honing from a good
+player~\cite{gelly2012go} to the best one with AlphaGo on one of the most highly
+combinatorial game wich is the game of Go entices us to use one for this
+problem. The Monte Carlo tree search~\cite{browne2012survey} algorithm will be
+used to fullfill this task.
+
+\section{Previous related works}
+\label{sec:previous_related_works}
 

### 2018-01-31 article/article.tex
(+) debut etat de l'art

@@
 \usepackage{abstract}
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
+% Set the "Abstract" text to bold
+\renewcommand{\abstractnamefont}{\normalfont\bfseries}
+% Set the abstract itself to small italic text
+\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
 
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
+% Allows customization of titles
+\usepackage{titlesec}
+% Roman numerals for the sections
+\renewcommand\thesection{\Roman{section}}
+% roman numerals for subsections
+\renewcommand\thesubsection{\roman{subsection}}
+% Change the look of the section titles
+\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
+% Change the look of the section titles
+\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{}
 
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-
+% Headers and footers
+\usepackage{fancyhdr}
+% All pages have headers and footers
+\pagestyle{fancy}
+% Blank out the default header
+\fancyhead{}
+% Blank out the default footer
+\fancyfoot{}
+% Custom header text
+\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
+% Custom footer text
+\fancyfoot[RO,LE]{\thepage}
 \usepackage{titling} % Customizing the title section
@@
 
-%----------------------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
 %	TITLE SECTION
-%----------------------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
 
-\setlength{\droptitle}{-4\baselineskip} % Move the title up
-
-\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
-\posttitle{\end{center}} % Article title closing formatting
+% Move the title up
+\setlength{\droptitle}{-4\baselineskip}
+% Article title formatting
+\pretitle{\begin{center}\Huge\bfseries}
+% Article title closing formatting
+\posttitle{\end{center}}
 
@@
 
-%----------------------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
 
@@
 
+The dynamic airspace configuration problem requires a model of the airspace. As
+done in~\cite{sergeeva2017dynamic} or~\cite{treimuth2016branch} the airspace
+is modelled via a graph. In those graphs, vertices represent elementary modules
+and an edge links two adjacent modules. In~\cite{treimuth2016branch}, to be
+able to produce a sequence of configurations, the graph is time dependent.
+An other way to do the modelization, is to use a constrainted set of
+configurations as in~\cite{gianazza2010forecasting} and~\cite{bloem2010dynamic}.
+This way any configuration will match specified requirements, which can be
+qualified hard constraints.
+
+In most cases the cost of a configuration is based on the workload. Each
+approach seems to give their own representation of the workload. For instance,
+~\cite{bedouet2016} determined workload density proportionally to the time
+spent by aircrafts in each sector. A simpler version~\cite{sergeeva2017dynamic}
+only uses the number of aircratfs. On the other hand, more complex method,
+involving many more inputs are also available, such as a neural network
+forecasting the workload~\cite{gianazza2010forecasting}. This is the method we
+are going to use in this paper.
+
 \bibliography{article}

### 2018-02-01 article/article.tex
(+) soft constraints

@@
 
+Other soft constraints appear to be relevant to have a better modelisation of
+the problem. For instance in~\cite{sergeeva2017dynamic} and~\cite{bedouet2016} a
+coordination cost is defined. It represent the surplus of work added by the
+number of flights travelling from a sector to an other. The shape of the
+resulting sector is considered, as the simpler is
+the shape, the easier it is to manage, therefore complex shapes are avoided,
+by introducing compactness in~\cite{jagare2013constraint} and balconies
+in~\cite{sergeeva2017dynamic}. To smooth the transition between two
+configurations, the work associated with the realocation of one or more modules
+is evaluated. This can be used in a cost function aimed to be minimized
+(in~\cite{bedouet2016}).
+
 \bibliography{article}

### 2018-02-01 article/article.tex
(-) fancy head
(~) some language fix

--- 2018-01-31 article/article.tex
@@
 % Blank out the default header
-\fancyhead{}
+% \fancyhead{}
 % Blank out the default footer
@@
 % Custom header text
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
+% \fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
 % Custom footer text
@@
 able to produce a sequence of configurations, the graph is time dependent.
-An other way to do the modelization, is to use a constrainted set of
+An other way to model the problem is to use a constrained set of
 configurations as in~\cite{gianazza2010forecasting} and~\cite{bloem2010dynamic}.
@@
 
+Other soft constraints appear to be relevant to have a better model of
+the problem. For instance in~\cite{sergeeva2017dynamic} and~\cite{bedouet2016} a
+coordination cost is defined. It represents the surplus of work added by
+flights travelling from one sector to an other. The shape of the
+resulting sector is considered, as the simpler is
+the shape, the easier it is to manage. Complex shapes are therefore avoided,
+thanks to the notion of compactness in~\cite{jagare2013constraint} and balconies
+in~\cite{sergeeva2017dynamic}. To smooth the transition between two
+configurations, the work associated with the reallocation of one or more modules
+is evaluated. This can be used in a cost function aimed to be minimized
+(in~\cite{bedouet2016}).
+
 \bibliography{article}

### 2018-02-02 article/article.tex
(~) aircraft has no plural

@@
 ~\cite{bedouet2016} determined workload density proportionally to the time
-spent by aircrafts in each sector. A simpler version~\cite{sergeeva2017dynamic}
-only uses the number of aircratfs. On the other hand, more complex method,
+spent by aircraft in each sector. A simpler version~\cite{sergeeva2017dynamic}
+only uses the number of aircraft. On the other hand, more complex method,
 involving many more inputs are also available, such as a neural network
@@
 configurations, the work associated with the reallocation of one or more modules
-is evaluated. This can be used in a cost function aimed to be minimized
+is evaluated. This can be used in a cost function aimed to be minimised
 (in~\cite{bedouet2016}).

### 2018-04-13 article/article.tex
(+) added mcts algo (not definitive)
(~) a bit more precise on the state of the art

@@
 \usepackage{hyperref}
+\usepackage{unicode-math}
+\usepackage{amsmath,amsthm,stmaryrd}
+\usepackage{algorithm,algpseudocode}
 
@@
 spent by aircraft in each sector. A simpler version~\cite{sergeeva2017dynamic}
-only uses the number of aircraft. On the other hand, more complex method,
-involving many more inputs are also available, such as a neural network
-forecasting the workload~\cite{gianazza2010forecasting}. This is the method we
-are going to use in this paper.
+only uses the number of aircraft. On the other hand, more complex methods,
+involving many more inputs are also available. For instance
+,~\cite{gianazza2010forecasting} used several indicators, such as sector
+volume, or vertical incoming flows in the next 15 and 60 minutes. Those
+indicators ease the training of a neural network forecasting the workload.
 
@@
 
+\section{Model}
+
+\section{Algorithm}
+\subsection{Monte Carlo tree search overview}
+\begin{algorithm}
+  \caption{General MCTS~\cite{browne2012survey}}
+  \begin{algorithmic}
+    \Procedure{MctsSearch}{$v_0$}
+    \While{within computational budget}
+    \State{} \(v_l \gets\) \Call{TreePolicy}{$v_0$}
+    \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$s(v_l)$}
+    \State{} \Call{Backup}{$v_l$, $\Delta$}
+    \EndWhile{}
+    \Return{\(a(\)\Call{BestChild}{$v_0$}\()\)}
+    \EndProcedure{}
+  \end{algorithmic}
+
+\end{algorithm}
+
 \bibliography{article}

### 2018-04-13 article/article.tex
(+) Added algorithms
(+) Continued article

@@
 \usepackage{algorithm,algpseudocode}
+\usepackage{paralist}
+
+\DeclareMathOperator{\argmax}{argmax}
 
@@
 only uses the number of aircraft. On the other hand, more complex methods,
-involving many more inputs are also available. For instance
-,~\cite{gianazza2010forecasting} used several indicators, such as sector
+involving many more inputs are also available. For
+instance,~\cite{gianazza2010forecasting} used several indicators, such as sector
 volume, or vertical incoming flows in the next 15 and 60 minutes. Those
@@
 
-\section{Model}
+\section{Algorithm}
 
-\section{Algorithm}
+The task of building an optimal sequence of partitions (lowering
+as much as possible the workload of each controller) is fulfilled by a
+stochastic tree search method, namely the Monte Carlo tree search. To assess the
+exactness of the results, an exact method (here A\(^*\)) is used.
+
 \subsection{Monte Carlo tree search overview}
+Monte-Carlo is a best-first search method using stochastic simulations. It is
+composed of four phases, namely
+\begin{compactenum}
+	\item selection,
+	\item expansion,
+	\item simulation,
+	\item backpropagation.
+\end{compactenum}
+Those phases are repeated until a stopping criterium is reached,
+see\ref{alg:gen_mcts} (memory usage or time).
 \begin{algorithm}
-  \caption{General MCTS~\cite{browne2012survey}}
+	\caption{General MCTS~\cite{browne2012survey}}\label{alg:gen_mcts}
 	\begin{algorithmic}
-    \Procedure{MctsSearch}{$v_0$}
+		\Procedure{MctsSearchTree}{$v_0$}
 		\While{within computational budget}
-    \State{} \(v_l \gets\) \Call{TreePolicy}{$v_0$}
-    \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$s(v_l)$}
-    \State{} \Call{Backup}{$v_l$, $\Delta$}
+		\State{} \([v^l_1, \dots, v^l_i] \gets\) \Call{TreePolicy}{$v_0$}
+		\State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v^l_i$}
+		\State{} \(r \gets\) \Call{Reward}{$\Delta$}
+		\State{} \Call{Backup}{$[v^l_1, \dots, v^l_i]$, $r$}
 		\EndWhile{}
-    \Return{\(a(\)\Call{BestChild}{$v_0$}\()\)}
 		\EndProcedure{}
 	\end{algorithmic}
+\end{algorithm}
 
+\subsection{UCT selection}
+\begin{algorithm}
+	\caption{Extending the search tree}
+	\begin{algorithmic}
+		\Function{TreePolicy}{$v$}
+		\State{} \([v_1, \dots, v_n] \gets\) \Call{Select}{$v$}
+		\If{$v_n$ is terminal}
+		\State{} \Return{$[v_1, \dots, v_n]$}
+		\Else{}
+		\State{} \([v_1, \dots, v_{n+1}] \gets\)
+		\Call{Expand}{$[v_1, \dots, v_n]$}
+		\State{} \Return{$[v_1, \dots, v_{n+1}]$}
+		\EndIf{}
+		\EndFunction{}
+	\end{algorithmic}
+	\begin{algorithmic}
+		\Function{Select}{$v, a$}
+		\If{$v$ is terminal or fully expanded}
+		\State{} \Return{$[v, a]$}
+		\Else{}
+		\State{} fav\(\gets\) \Call{BestChild}{$v$}
+		\State{} \Return{\Call{Select}{fav, $[v, a]$}}
+		\EndIf{}
+		\EndFunction{}
+	\end{algorithmic}
+	\begin{algorithmic}
+		\Function{BestChild}{$v$}
+		\State{} \Return{$\argmax\{\Call{UCB}{v'} |
+		v' \text{\,children of\,} v\}$}
+		\EndFunction{}
+	\end{algorithmic}
 \end{algorithm}
+
+\section{Model}
 

### 2018-04-16 article/article.tex
(+) small par on mcts

@@
 
-\subsection{Monte Carlo tree search overview}
-Monte-Carlo is a best-first search method using stochastic simulations. It is
-composed of four phases, namely
+\subsection{Monte Carlo tree search}
+\subsubsection{Overview}
+Monte-Carlo is a best-first search method using stochastic simulations. The
+algorithm actually uses two trees, an underlying tree associated to the model
+(e.g.\ a game tree) and a search tree. The latter is built incrementally by the
+algorithm the following way, at each step, starting from the root, the most
+promising nodes towards the bottom of the tree are selected. Once a leaf of the
+search tree is reached, a new node coming from the model tree is added to the
+leaf. A simulation is then run from this newly added node to evaluate the
+outcome of a path starting from it. The result of the simulation is eventually
+backpropagated to the consecutive fathers of the new node.
 \begin{compactenum}

### 2018-04-16 article/article.tex
(+) notations

--- 2018-04-13 article/article.tex
@@
 
+In this section, we discuss the model established to approach this configuration
+problem. First we defined a structure for a controled sector and all possible
+transitions through time. Then a workload model is defined. And finally we
+determined a cost per partition and a cost function aimed to be minimized.
+
+\subsection{State}
+
+\subsection{workload}
+
+$p_{low}$
+$p_{normal}$
+$p_{high}$
+
+\subsection{Partition cost and cost function}
+
+In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
+to be defined. This cost depend on the workload in each the sector. We give the
+following intermediate definitions to represent a high (respectively normal
+and low) cost for the partition :
+\begin{itemize}
+    \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
+    \item $c_=(P, t) = \sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$
+    \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
+\end{itemize}
+with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
+if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
+the two other, and 0 otherwise.
+
+With
+
+\section*{Notation}
+
+\begin{itemize}
+    \item $m$ : elementary module
+    \item $S$ : sector control position, group of modules
+    \item $P$ : partition of the airspace, group of sectors
+    \item $t$ : current time
+    \item $C(P, t)$ : Cost of a given partition $P$ at time $t$
+    \item $s$ : state, in this article a partition $P$ and a time $t$
+    \item $h(s)$ : heuristic that gives for a state $s$ an estimation of the cost
+    \item $\mu_{F, a}$ : mean of all reward comming from node $F$ with action $a$
+    \item $T_F$ : number of simuation where the final path include the node $F$
+    \item $T_{F, a}$ : number of simuation where the final path include the node $F$ with action $a$
+\end{itemize}
+
 \bibliography{article}

### 2018-04-16 article/article.tex
(+) notations

@@
 
-\section*{Notation}
+\section*{Notations}
 

### 2018-04-16 article/article.tex
(+) continuing

@@
 to be defined. This cost depend on the workload in each the sector. We give the
-following intermediate definitions to represent a high (respectively normal
+following definitions~\cite{Ferrari2017} to represent a high (respectively normal
 and low) cost for the partition :
@@
 if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
-the two other, and 0 otherwise.
+the two others, and 0 otherwise.
 
-With
+It is now possible to assign a cost to each partition. This cost is linear
+regarding
 

### 2018-04-16 article/article.tex
(+) Added par on selection and expansion

--- 2018-04-16 article/article.tex
@@
 backpropagated to the consecutive fathers of the new node.
+
+The above mentioned steps are denoted
 \begin{compactenum}
-    \item selection,
-    \item expansion,
-    \item simulation,
-    \item backpropagation.
+	\item selection (or tree walk): choosing successively most promising nodes,
+	\item expansion: adding new nodes to the search tree,
+	\item simulation (or random walk): choosing quickly actions from the model
+		tree,
+	\item backpropagation (or backup): applying the result of the simulation to
+		the search tree.
 \end{compactenum}
-Those phases are repeated until a stopping criterium is reached,
-see\ref{alg:gen_mcts} (memory usage or time).
+Those phases are repeated until a stopping criterium (e.g.\ memory or time) is
+reached, resulting in algorithm~\ref{alg:gen_mcts}.
 \begin{algorithm}
@@
 		\While{within computational budget}
-        \State{} \([v^l_1, \dots, v^l_i] \gets\) \Call{TreePolicy}{$v_0$}
-        \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v^l_i$}
-        \State{} \(r \gets\) \Call{Reward}{$\Delta$}
-        \State{} \Call{Backup}{$[v^l_1, \dots, v^l_i]$, $r$}
+		\State{} \([v_0, \dots, v_n] \gets\) \Call{TreePolicy}{$v_0$}
+		\State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v_n$}
+		\State{} \Call{Backup}{$[v_0, \dots, v_n]$, $\Delta$}
 		\EndWhile{}
@@
 
-\subsection{UCT selection}
+\subsubsection{Extending the search tree}
+\paragraph{Selection}
+The aim of the selection is to choose from a set of actions, the most promising
+one\footnote{i.e.\ possibly leading to the best evaluation} according to a
+\emph{tree policy}. This type of problem is addressed by bandits methods where
+a multi armed bandit aims to always activate the slot machine giving the highest
+reward.
+
+The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
+It uses the UCB for Upper Confidence Bound as tree policy (introduced
+in~\cite{kocsis2006bandit}). Let \(F\) be the node from which an action must be
+selected to go deeper in the tree, \(a\) an action available from \(F\),
+\(\mu_{F,a}\) the expected outcome from the node obtained by applying action
+\(a\) on \(F\), \(T_F\) the number of simulations carried out from node \(F\)
+(which includes any simulation from nodes in any subtree of \(F\)) and
+\(T_{F,a}\) the number of simulations carried out from the node resulting from
+the action \(a\) applied to node \(F\), \(\beta\) a constant. The selected node
+is thus the one maximising
+\begin{equation}
+	\text{UCB} = \mu_{F,a} + 2 \beta \sqrt{\frac{2 \log T_f}{T_{F,a}}}
+\end{equation}
+
+\paragraph{Expansion}
+The above step is carried out only if, considering a node \(F\), a simulation
+has been run through each child available from \(F\). In the case where one or
+more children haven't been explored yet, one of the unexplored children is
+chosen and a simulation is started from it.
 \begin{algorithm}
@@
 
-In this section, we discuss the model established to approach this configuration
-problem. First we defined a structure for a controled sector and all possible
-transitions through time. Then a workload model is defined. And finally we
-determined a cost per partition and a cost function aimed to be minimized.
-
-\subsection{State}
-
-\subsection{workload}
-
-$p_{low}$
-$p_{normal}$
-$p_{high}$
-
-\subsection{Partition cost and cost function}
-
-In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
-to be defined. This cost depend on the workload in each the sector. We give the
-following intermediate definitions to represent a high (respectively normal
-and low) cost for the partition :
-\begin{itemize}
-    \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
-    \item $c_=(P, t) = \sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$
-    \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
-\end{itemize}
-with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
-if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
-the two other, and 0 otherwise.
-
-With
-
-\section*{Notations}
-
-\begin{itemize}
-    \item $m$ : elementary module
-    \item $S$ : sector control position, group of modules
-    \item $P$ : partition of the airspace, group of sectors
-    \item $t$ : current time
-    \item $C(P, t)$ : Cost of a given partition $P$ at time $t$
-    \item $s$ : state, in this article a partition $P$ and a time $t$
-    \item $h(s)$ : heuristic that gives for a state $s$ an estimation of the cost
-    \item $\mu_{F, a}$ : mean of all reward comming from node $F$ with action $a$
-    \item $T_F$ : number of simuation where the final path include the node $F$
-    \item $T_{F, a}$ : number of simuation where the final path include the node $F$ with action $a$
-\end{itemize}
-
 \bibliography{article}

### 2018-04-16 article/article.tex

@@
 
+In this section, we discuss the model established to approach this configuration
+problem. First we defined a structure for a controled sector and all possible
+transitions through time. Then a workload model is defined. And finally we
+determined a cost per partition and a cost function aimed to be minimized.
+
+\subsection{State}
+
+\subsection{workload}
+
+$p_{low}$
+$p_{normal}$
+$p_{high}$
+
+\subsection{Partition cost and cost function}
+
+In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
+to be defined. This cost depend on the workload in each the sector. We give the
+following definitions~\cite{Ferrari2017} to represent a high
+(respectively normal and low) cost for the partition:
+\begin{itemize}
+  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
+  \item $c_=(P, t) = \sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$
+  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
+\end{itemize}
+with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
+if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
+the two others, and 0 otherwise.
+
+It is now possible to assign a cost to each partition. This cost is linear
+regarding
+
+\section*{Notations}
+
+\begin{itemize}
+  \item $m$: elementary module
+  \item $S$: sector control position, group of modules
+  \item $P$: partition of the airspace, group of sectors
+  \item $t$: current time
+  \item $C(P, t)$: cost of a given partition $P$ at time $t$
+  \item $s$: state, in this article a partition $P$ and a time $t$
+  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
+  \item $\mu_{F, a}$: mean of all reward comming from node $F$ with action $a$
+  \item $T_F$: number of simuation where the final path include the node $F$
+  \item $T_{F, a}$: number of simuation where the final path include the
+    node $F$ with action $a$
+\end{itemize}
+
 \bibliography{article}

### 2018-04-18 article/article.tex
(~) moved notations to head
(+) Rewritten a bit algorithms, new notations

@@
 \DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator{\cons}{::}
 
@@
 
+\section*{Notations}
+\begin{itemize}
+  \item $m$: elementary module
+  \item $S$: sector control position, group of modules
+  \item $P$: partition of the airspace, group of sectors
+  \item $t$: current time
+  \item $C(P, t)$: cost of a given partition $P$ at time $t$
+  \item $s$: state, in this article a partition $P$ and a time $t$
+  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
+  \item $\mu_{v, a}$: mean of all reward comming from node $v$ with action $a$
+  \item $T_v$: number of simulation where the final path include the node $v$
+  \item $T_{v, a}$: number of simulation where the final path include the
+    node $F$ with action $a$
+  \item \(\mathcal{A}(v)\) the available actions from a node \(v\)
+  \item \(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\) the
+    production function, giving the resulting state of action \(a\) applied on a
+    state \(s\).
+\end{itemize}
+
 \section{Algorithm}
@@
 \subsubsection{Extending the search tree}
-\paragraph{Selection}
+\paragraph{Selection~\ref{alg:tree_pol}}
 The aim of the selection is to choose from a set of actions, the most promising
@@
 It uses the UCB for Upper Confidence Bound as tree policy (introduced
-in~\cite{kocsis2006bandit}). Let \(F\) be the node from which an action must be
-selected to go deeper in the tree, \(a\) an action available from \(F\),
-\(\mu_{F,a}\) the expected outcome from the node obtained by applying action
-\(a\) on \(F\), \(T_F\) the number of simulations carried out from node \(F\)
-(which includes any simulation from nodes in any subtree of \(F\)) and
-\(T_{F,a}\) the number of simulations carried out from the node resulting from
-the action \(a\) applied to node \(F\), \(\beta\) a constant. The selected node
+in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
+selected to go deeper in the tree, \(a\) an action available from \(v\),
+\(\mu_{v,a}\) the expected outcome from the node obtained by applying action
+\(a\) on \(v\), \(T_v\) the number of simulations carried out from node \(v\)
+(which includes any simulation from nodes in any subtree of \(v\)) and
+\(T_{v,a}\) the number of simulations carried out from the node resulting from
+the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
 is thus the one maximising
 \begin{equation}
-  \text{UCB} = \mu_{F,a} + 2 \beta \sqrt{\frac{2 \log T_f}{T_{F,a}}}
+  \text{UCB} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
 \end{equation}
@@
 \begin{algorithm}
-  \caption{Extending the search tree}
+  \caption{UCT algorithm}\label{alg:tree_pol}
   \begin{algorithmic}
-    \Function{TreePolicy}{$v$}
-    \State{} \([v_1, \dots, v_n] \gets\) \Call{Select}{$v$}
-    \If{$v_n$ is terminal}
-    \State{} \Return{$[v_1, \dots, v_n]$}
+    \Function{TreePolicy}{path of nodes $\pi$}
+    \State{} $v \gets$ head of $\pi$
+    \State{} \Call{Deploy}{$v$}
+    \If{$v$ is terminal}
+    \State{}\Return{$\pi$}
     \Else{}
-    \State{} \([v_1, \dots, v_{n+1}] \gets\)
-    \Call{Expand}{$[v_1, \dots, v_n]$}
-    \State{} \Return{$[v_1, \dots, v_{n+1}]$}
+    \If{$v$ is not fully expanded}
+    \State{}\Return{\Call{Expand}{$v$} $\cons \pi$}
+    \Else{}
+    \State $f \gets$ \Call{BestChild}{$v$}
+    \State{}\Return{%
+      \Call{TreePolicy}{$f \cons \pi$}
+    }
     \EndIf{}
-    \EndFunction{}
-  \end{algorithmic}
-  \begin{algorithmic}
-    \Function{Select}{$v, a$}
-    \If{$v$ is terminal or fully expanded}
-    \State{} \Return{$[v, a]$}
-    \Else{}
-    \State{} fav\(\gets\) \Call{BestChild}{$v$}
-    \State{} \Return{\Call{Select}{fav, $[v, a]$}}
     \EndIf{}
@@
     v' \text{\,children of\,} v\}$}
+    \EndFunction{}
+  \end{algorithmic}
+\end{algorithm}
+\begin{algorithm}
+  \caption{%
+    Expansion (\(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\)
+    is the transition function)
+  }
+  \begin{algorithmic}
+    \Function{Expand}{node $v$}
+    \State{} $\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
+    \State{}\Return{random element from $\ell$}
     \EndFunction{}
@@
 
-\section*{Notations}
-
-\begin{itemize}
-  \item $m$: elementary module
-  \item $S$: sector control position, group of modules
-  \item $P$: partition of the airspace, group of sectors
-  \item $t$: current time
-  \item $C(P, t)$: cost of a given partition $P$ at time $t$
-  \item $s$: state, in this article a partition $P$ and a time $t$
-  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
-  \item $\mu_{F, a}$: mean of all reward comming from node $F$ with action $a$
-  \item $T_F$: number of simuation where the final path include the node $F$
-  \item $T_{F, a}$: number of simuation where the final path include the
-    node $F$ with action $a$
-\end{itemize}
-
 \bibliography{article}

### 2018-04-18 article/article.tex
(+) Continued

@@
 \begin{equation}
-  \text{UCB} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
+  \text{UCB1} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
 \end{equation}
 
-\paragraph{Expansion}
-The above step is carried out only if, considering a node \(F\), a simulation
-has been run through each child available from \(F\). In the case where one or
-more children haven't been explored yet, one of the unexplored children is
-chosen and a simulation is started from it.
+While the previous equation is well suited for two players games, it can be
+tweaked to improve its efficience in one player games or sequencing problems. An
+alternative using the standard deviation is proposed in~\cite{sebag2010fuse}
+called the \emph{UCB1-tuned} equation
+\begin{equation}
+  \mu_{v,a} +
+  \sqrt{%
+    \frac{\beta\log T_v}{t_{v,a}}
+    \min\left( \frac{1}{4}, \sigma^2_{v,a} +
+    \sqrt{%
+      \frac{2 \log T_v}{T_{v,a}}
+    }\right)
+  }
+\end{equation}
 \begin{algorithm}
@@
     \Function{BestChild}{$v$}
-    \State{} \Return{$\argmax\{\Call{UCB}{v'} |
+    \State{} \Return{$\argmax\{\Call{UCB1-tuned}{v'} |
     v' \text{\,children of\,} v\}$}
@@
 \end{algorithm}
+
+\paragraph{Expansion~\ref{alg:expansion}}
+The above step is carried out only if, considering a node \(v\), a simulation
+has been run through each child available from \(v\). In the case where one or
+more children haven't been explored yet, one of the unexplored children is
+chosen and a simulation is started from it.
 \begin{algorithm}
@@
     is the transition function)
-  }
+  }\label{alg:expansion}
   \begin{algorithmic}
@@
 
+\subsubsection{Simulation and backpropagation}
+\paragraph{Simulation}
+Once a node has been expanded, random nodes are chosen successively until a
+terminal state is found. The cost of the resulting path, from the root to the
+node is then evaluated and backpropagated to all the ancestors of the expanded
+node.
+
+\subparagraph{Heuristic} One might want to bias the randomness while choosing
+nodes. This would imply using a heuristic which has to be able to discriminate
+a node between its siblings.
+
+\paragraph{Backpropagation}
+The backpropagation consists in updating the values required to carry out the
+tree policy. The values must be updated incrementally since the backpropagation
+function has the current value of the parameters and the result of the
+simulation as parameters. Thus, for UCB1 equation, the expected reward (mean)
+and the total count are updated this way
+\begin{algorithm}
+  \caption{UCB1 backpropagation}
+  \begin{algorithmic}
+    \Procedure{Backpropagate}{path $\pi$, $r$}
+    \State{} \(\delta \gets\ r - \mu_v\)
+    \State{} \(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
+    \State{} \(T_v \gets T_v + 1\)
+    \EndProcedure{}
+  \end{algorithmic}
+\end{algorithm}
+
+For the UCB1-tuned, the standard deviation has to be computet. It results in
+algorithm~\ref{alg:backpropagate-tuned}
+\begin{algorithm}
+  \caption{UCB1-tuned backpropagation}\label{alg:backpropagate-tuned}
+\end{algorithm}
 \section{Model}

### 2018-04-18 article/article.tex
(+) almost finished mcts

@@
 problem, for instance via genetic algorithms in~\cite{sergeeva2017dynamic},
-constraint local search in~\cite{jagare2013constraint}, integer linear
+constraint local search in~\cite{jagare2013airspace}, integer linear
 programming in~\cite{treimuth2016branch} or dynamic programming
@@
 sochastic tree search algorithm has been honing from a good
-player~\cite{gelly2012go} to the best one with AlphaGo on one of the most highly
-combinatorial game wich is the game of Go entices us to use one for this
+player~\cite{gelly2012grand} to the best one with AlphaGo on one of the most
+highly combinatorial game wich is the game of Go entices us to use one for this
 problem. The Monte Carlo tree search~\cite{browne2012survey} algorithm will be
@@
 approach seems to give their own representation of the workload. For instance,
-~\cite{bedouet2016} determined workload density proportionally to the time
+~\cite{bedouet2016towards} determined workload density proportionally to the time
 spent by aircraft in each sector. A simpler version~\cite{sergeeva2017dynamic}
@@
 Other soft constraints appear to be relevant to have a better model of
-the problem. For instance in~\cite{sergeeva2017dynamic} and~\cite{bedouet2016} a
+the problem. For instance in~\cite{sergeeva2017dynamic}
+and~\cite{bedouet2016towards} a
 coordination cost is defined. It represents the surplus of work added by
@@
 the shape, the easier it is to manage. Complex shapes are therefore avoided,
-thanks to the notion of compactness in~\cite{jagare2013constraint} and balconies
+thanks to the notion of compactness in~\cite{jagare2013airspace} and balconies
 in~\cite{sergeeva2017dynamic}. To smooth the transition between two
@@
 is evaluated. This can be used in a cost function aimed to be minimised
-(in~\cite{bedouet2016}).
+(in~\cite{bedouet2016towards}).
 
@@
     \State{} $v \gets$ head of $\pi$
-    \State{} \Call{Deploy}{$v$}
     \If{$v$ is terminal}
@@
     \If{$v$ is not fully expanded}
-    \State{}\Return{\Call{Expand}{$v$} $\cons \pi$}
+    \State{}\Return{\Call{Expand}{$v$} $\cup \pi$}
     \Else{}
-    \State $f \gets$ \Call{BestChild}{$v$}
+    \State{}$f \gets$ \Call{BestChild}{$v$}
     \State{}\Return{%
-      \Call{TreePolicy}{$f \cons \pi$}
+      \Call{TreePolicy}{$f \cup \pi$}
     }
@@
     \Function{BestChild}{$v$}
-    \State{} \Return{$\argmax\{\Call{UCB1-tuned}{v'} |
+    \State{}\Return{$\argmax\{\Call{UCB}{v'} |
     v' \text{\,children of\,} v\}$}
@@
     \Function{Expand}{node $v$}
-    \State{} $\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
+    \State{}$\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
     \State{}\Return{random element from $\ell$}
@@
   \begin{algorithmic}
-    \Procedure{Backpropagate}{path $\pi$, $r$}
-    \State{} \(\delta \gets\ r - \mu_v\)
-    \State{} \(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
-    \State{} \(T_v \gets T_v + 1\)
+    \Procedure{Backpropagate}{$v$, $r$}
+    \State{}\(\delta \gets\ r - \mu_v\)
+    \State{}\(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
+    \State{}\(T_v \gets T_v + 1\)
     \EndProcedure{}
@@
 
-For the UCB1-tuned, the standard deviation has to be computet. It results in
-algorithm~\ref{alg:backpropagate-tuned}
+For the UCB1-tuned, the standard deviation has to be computed. It results in
+algorithm~\ref{alg:backpropagate-tuned} which introduces the value \(m_v\)
+associated to a node \(v\) to compute the standard deviation via the formula
+\begin{equation}
+  \sigma^2_v = \frac{m_v}{T_v}.
+\end{equation}
 \begin{algorithm}
   \caption{UCB1-tuned backpropagation}\label{alg:backpropagate-tuned}
+  \begin{algorithmic}
+    \Procedure{Backpropagate}{$v, r$}
+    \State{}\(\delta \gets r - \mu_v\)
+    \State{}\(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
+    \State{}\(\delta' \gets r - \mu_v\)
+    \State{}\(m_v \gets m_v + \delta \delta'\)
+    \State{}\(T_v \gets T_v + 1\)
+    \EndProcedure{}
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Sequence building}
+\subsubsection{Node selection}\label{sssec:node_selection}
+The building of the final sequence of actions differs from the selection on the
+aim, the selection chooses the most promising actions while the final sequence
+must be composed of the best. This means that the final path build won't use the
+UCB equation. Chaslot \textit{et al.}~\cite{browne2012survey} propose several
+methods to select nodes,
+\begin{itemize}
+  \item max-child: select the child with highest reward;
+  \item robust child: select the most visited root child;
+  \item  secure child: select the child maximising a lower confidence bound.
+\end{itemize}
+
+\subsubsection{Iterative pathfinding}
+The path is built iteratively by calling successive Monte Carlo tree searches.
+Say an MCTS has been called on a node \(u\). Once a stopping criterium is
+matched, a node \(v\) among the ones reachable from \(u\) is selected via one of
+the previously mentioned policies. Then an mcts is called back with node \(u\)
+as the root node. The path is composed of the successive roots. The algorithm is
+described at~\ref{alg:path_building}.
+\begin{algorithm}
+  \caption{%
+    Path building. The {\sc NextNode} (here max-child) function is one among
+    those in~\ref{sssec:node_selection}.
+  }\label{alg:path_building}
+  \begin{algorithmic}
+    \Function{MctsSearch}{$u_0, n$}
+    \State{}\(u \gets u_0\)
+    \State{}\(\pi \gets \{u\}\)
+    \For{$i = 1$ to $n$}
+      \State{}\Call{MctsSearchTree}{$u$}
+      \State{}\(u \gets\)\Call{NextNode}{$u$}
+      \State{}\(\pi \gets \pi \cup \{u\}\)
+    \EndFor{}
+    \State{}\Return{$\pi$}
+    \EndFunction{}
+    \Function{NextNode}{$u$}
+    \State{}\Return{$\argmax\{\mu_{u,a} | a \in \mathcal{A}(v)\}$}
+    \EndFunction{}
+  \end{algorithmic}
 \end{algorithm}

### 2018-04-18 script/test.sh
(+) script for results

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+for i in 1 10 50 100 250 500 1000 5000 10000; do
+    ../airmcts.native -scenario ../data/scen4.json -horizon 20 -nsteps 20 -timeperstep 1 -expvexp $i
+done

### 2018-04-20 article/article.tex
(+) partition cost and transition cost

@@
 
+\subsubsection{Partition cost}
+
 In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
@@
   \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
-  \item $c_=(P, t) = \sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$
+  \item $c_=(P, t) = 1 / (\sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$)
   \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
@@
 It is now possible to assign a cost to each partition. This cost is linear
-regarding
+regarding $c_+$, $c_=$, $c_-$ and $n$ (cardinal of the partition $P$). The
+partition cost is defined as follow:
+$$
+C(P, t) = \alpha c_+ + \beta c_= + \gamma c_- +\lambda n
+$$
+
+The parameters $\alpha$, $\beta$, $\gamma$ and $\lambda$ (all positive)
+determine a priority on which parameter to optimize. For instance, a high
+value $\alpha$ help to
+minimize $c_+$, hence the overall number of sectors with too much tafic.
+In a real application, it maybe interesting to order those parmaeters as
+follow : $\alpha > \lambda > \beta > \gamma$.
+
+\subsubsection{Transition cost}
+
+In an operational context, each reconfiguration raise the workload for
+controllers. This trend needs to be considered and a transition cost needs to
+be defined for this matter. A transition is (TODO), and the transition cost is
+then:
+$$
+C_{tr}(P_1, P_2) =
+\left\{
+\begin{array}{l}
+    0 \qquad if \qquad P_1 = P_2 \\
+    1 \qquad otherwise
+\end{array}
+\right.
+$$
 

### 2018-04-20 article/article.tex
(+) continued...

--- 2018-04-18 article/article.tex
@@
 \end{algorithm}
+\subsection{A\(^*\)}
+To evaluate the exactness of the paths given by the MCTS algoithm, the A\(^*\)
+algorithm is used. A\(^*\) is an exact depth first search pathfinding algorithm
+which uses a heuristic function to guide its search. The algorithm is given
+in~\ref{alg:astar} where \(u_0\) is the initial state, \(T\) the set of termial
+nodes, \(h\) a heuristic function estimating the cost from a state \(u\) given
+as argument to a final state, \(k \colon \mathcal{S}^2 \to \mathbb{R}\). The
+function \(\mathup{first(G)}\) returns the first element of \(G\) and
+\(f\mathup{-insert}(G, v)\) inserts \(v\) in \(G\) ordering first by \(f\)
+increasing then by \(g\) decreasing.
+\begin{algorithm}
+  \caption{A\(^*\) algorithm~\cite{alliotschiex2002ia&it}}\label{alg:astar}
+  \begin{algorithmic}
+    \Procedure{A$^*$}{$u_0$}
+    \State{}\(G \gets u_0; D \gets \emptyset; g(u_0) \gets 0; f(u_0) \gets 0\)
+    \State{}\(\mathup{father}(u_0) \gets \emptyset\)
+    \While{$G \neq \emptyset$}
+    \State{}\(u \gets \mathup{first}(G); G \gets G \backslash \{u\}\)
+    \State{}\(D \gets D \cup \{u\}\)
+    \If{$u \in T$}
+    \State{}\Return{father}
+    \EndIf{}
+    \For{$v$ in childen of $u$}
+    \If{$v \notin D \cup G$ or $[g(v) > g(u) + k(u, v)]$}
+    \State{}\(g(v) \gets g(u) + k(u, v)\)
+    \State{}\(f(v) \gets g(v) + h(v)\)
+    \State{}\(\mathup{father}(v) \gets u\)
+    \State{}\(f\mathup{-insert}(G, v)\)
+    \EndIf{}
+    \EndFor{}
+    \EndWhile{}
+    \EndProcedure{}
+  \end{algorithmic}
+\end{algorithm}
 \section{Model}
@@
 to be defined. This cost depend on the workload in each the sector. We give the
-following definitions~\cite{Ferrari2017} to represent a high
+following definitions~\cite{ferrari2017} to represent a high
 (respectively normal and low) cost for the partition:

### 2018-04-20 article/article.tex
(+) objective function and cost function
+ smaller subsubsection and not bold

@@
 \titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{}
+\titleformat{\subsubsection}[block]{\normal}{\thesubsection.}{1em}{}
 
@@
 partition cost is defined as follow:
-$$
+\begin{equation}
 C(P, t) = \alpha c_+ + \beta c_= + \gamma c_- +\lambda n  
-$$
+\end{equation}
 
@@
 In a real application, it maybe interesting to order those parmaeters as
-follow : $\alpha > \lambda > \beta > \gamma$.
+follow : $\alpha > \gamma > \beta > \lambda$.
 
@@
 then:
-$$
+\begin{equation}
   C_{tr}(P_1, P_2) =
@@
 \right.
-$$
+\end{equation}
+
+\subsection{Objective function}
+
+With a cost defined for partition and transition, it is now possible to
+aggregate everything in order to build a cost function over an entire
+path. For a path ($path = [P_0, ..., P_n]$) with a time from $t_0$ to $t_n$,
+the cost function is given by:
+\begin{equation}
+  \begin{split}
+  f(path) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\ \theta \times C_{tr}(P_{i-1}, P_i)]   
+  \end{split}
+\end{equation}
+
+with $\theta > 0$ a parameter to determine.
+This is the imediate function to minimize in the A$^*$ algorithm.
+
+The Monte Carlo Tree search algorithm maximize a objective function. This function
+represent a potential reward over a path, i.e. a list of partitions over time.
+Given the loss function defined previously, the target function can be
+construct as:
+\begin{equation}
+  \psi(path) = \frac{1}{f(path)}
+\end{equation}
 

### 2018-04-20 article/article.tex
(-) no more custom commands
(+) subsubsub
(~) Differentiated notations
Algorithms systematically use nodes whereas equations use states.

--- 2018-04-20 article/article.tex
@@
 \DeclareMathOperator{\argmax}{argmax}
-\DeclareMathOperator{\cons}{::}
+\DeclareMathOperator{\first}{first}
+\DeclareMathOperator{\state}{state}
 
@@
 \maketitle
+
+\section*{Notations}
+\begin{itemize}
+  \item $m$: elementary module;
+  \item $S$: sector control position, group of modules;
+  \item $P$: partition of the airspace, group of sectors;
+  \item $t$: current time;
+  \item $C(P, t)$: cost of a given partition $P$ at time $t$;
+  \item \(\mathcal{S}\): the set of available states;
+  \item \(\mathcal{A}\): the set of all possible actions;
+  \item \(\mathcal{A}(s)\): the available actions from a state \(s\);
+  \item $s$: state;
+  \item \(f \colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\):
+    transition function giving the resulting states of action \(a\) executed
+    from a state \(s\).
+  \item $h \colon \mathcal{S} \to \mathbb{R}^+$: heuristic estimating the cost
+    from a state \(s \in \mathcal{S}\) to a final state;
+  \item \(\mathcal{N}\): the set of nodes
+  \item \(u, v \in \mathcal{N}^2\): nodes
+  \item \(Q \colon \mathcal{N} \to \mathbb{R}^+\): the reward expectancy of a
+    node;
+  \item \(N \colon \mathcal{N} \to \mathbb{N}\): number of simulations running
+    through or from a node;
+  \item \(\first(\ell)\): returns the first element of \(\ell\).
+\end{itemize}
 
@@
 
-\section*{Notations}
-\begin{itemize}
-  \item $m$: elementary module
-  \item $S$: sector control position, group of modules
-  \item $P$: partition of the airspace, group of sectors
-  \item $t$: current time
-  \item $C(P, t)$: cost of a given partition $P$ at time $t$
-  \item $s$: state, in this article a partition $P$ and a time $t$
-  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
-  \item $\mu_{v, a}$: mean of all reward comming from node $v$ with action $a$
-  \item $T_v$: number of simulation where the final path include the node $v$
-  \item $T_{v, a}$: number of simulation where the final path include the
-    node $F$ with action $a$
-  \item \(\mathcal{A}(v)\) the available actions from a node \(v\)
-  \item \(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\) the
-    production function, giving the resulting state of action \(a\) applied on a
-    state \(s\).
-\end{itemize}
-
 \section{Algorithm}
@@
 outcome of a path starting from it. The result of the simulation is eventually
-backpropagated to the consecutive fathers of the new node.
+backpropagated to the consecutive parents of the new node.
 
@@
 Those phases are repeated until a stopping criterium (e.g.\ memory or time) is
-reached, resulting in algorithm~\ref{alg:gen_mcts}.
+reached, resulting in algorithm~\ref{alg:gen_mcts} where the tree policy
+aggregates phases 1 and 2 to create a new node of the search tree and the
+default policy gives an evaluation of the newly added node.
 \begin{algorithm}
@@
 \subsubsection{Extending the search tree}
-\paragraph{Selection~\ref{alg:tree_pol}}
+\paragraph{Selection (algorithm~\ref{alg:tree_pol})}
 The aim of the selection is to choose from a set of actions, the most promising
@@
 The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
-It uses the UCB for Upper Confidence Bound as tree policy (introduced
+It uses the UCB (Upper Confidence Bound) as tree policy (introduced
 in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
@@
 the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
-is thus the one maximising
+is thus the one maximising the \emph{UCB1} equation
 \begin{equation}
-  \text{UCB1} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
+  \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
 \end{equation}
@@
 tweaked to improve its efficience in one player games or sequencing problems. An
-alternative using the standard deviation is proposed in~\cite{sebag2010fuse}
-called the \emph{UCB1-tuned} equation
+alternative using the standard deviation of the outcome of the simulations is
+proposed in~\cite{sebag2010fuse} called the \emph{UCB1-tuned} equation
 \begin{equation}
@@
   \begin{algorithmic}
-    \Function{TreePolicy}{path of nodes $\pi$}
-    \State{} $v \gets$ head of $\pi$
+    \Function{TreePolicy}{$\pi$}
+    \State{}\(v \gets \first(\pi)\)
     \If{$v$ is terminal}
@@
     \Function{BestChild}{$v$}
-    \State{}\Return{$\argmax\{\Call{UCB}{v'} |
-    v' \text{\,children of\,} v\}$}
+    \State{}\Return{$\argmax\{\Call{UCB}{v'} | v' \text{\,children of\,} v \}$}
     \EndFunction{}
@@
 
-\paragraph{Expansion~\ref{alg:expansion}}
+\paragraph{Expansion (algorithm~\ref{alg:expansion})}
 The above step is carried out only if, considering a node \(v\), a simulation
@@
 \begin{algorithm}
-  \caption{%
-    Expansion (\(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\)
-    is the transition function)
-  }\label{alg:expansion}
+  \caption{Expansion}\label{alg:expansion}
   \begin{algorithmic}
-    \Function{Expand}{node $v$}
-    \State{}$\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
+    \Function{Expand}{$v$}
+    \State{}$\ell \gets \{ v | v \text{\,children of\,} u, N(v) = 0 \}$
     \State{}\Return{random element from $\ell$}
@@
     \State{}\(\delta \gets\ r - \mu_v\)
-    \State{}\(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
-    \State{}\(T_v \gets T_v + 1\)
+    \State{}\(Q(v) \gets Q(v) + \frac{\delta}{N(v) + 1}\)
+    \State{}\(N(v) \gets N(v) + 1\)
     \EndProcedure{}
@@
 For the UCB1-tuned, the standard deviation has to be computed. It results in
-algorithm~\ref{alg:backpropagate-tuned} which introduces the value \(m_v\)
+algorithm~\ref{alg:backpropagate-tuned} which introduces the value \(m_2(v)\)
 associated to a node \(v\) to compute the standard deviation via the formula
 \begin{equation}
-  \sigma^2_v = \frac{m_v}{T_v}.
+  \sigma^2_v = \frac{m_2(v)}{N(v)}.
 \end{equation}
@@
     \Procedure{Backpropagate}{$v, r$}
-    \State{}\(\delta \gets r - \mu_v\)
-    \State{}\(\mu_v \gets \mu_v + \frac{\delta}{T_v + 1}\)
-    \State{}\(\delta' \gets r - \mu_v\)
-    \State{}\(m_v \gets m_v + \delta \delta'\)
-    \State{}\(T_v \gets T_v + 1\)
+    \State{}\(\delta \gets r - Q(v)\)
+    \State{}\(Q(v) \gets Q(v) + \frac{\delta}{N(v) + 1}\)
+    \State{}\(\delta' \gets r - Q(v)\)
+    \State{}\(m_2(v) \gets m_2(v) + \delta \delta'\)
+    \State{}\(N(v) \gets N(v) + 1\)
     \EndProcedure{}
@@
 methods to select nodes,
-\begin{itemize}
+\begin{compactitem}
   \item max-child: select the child with highest reward;
@@
   \item secure child: select the child maximising a lower confidence bound.
-\end{itemize}
+\end{compactitem}
 
@@
     \Function{NextNode}{$u$}
-    \State{}\Return{$\argmax\{\mu_{u,a} | a \in \mathcal{A}(v)\}$}
+    \State{}\Return{$\argmax\{Q(v) | v \text{\,children of\,} u\}$}
     \EndFunction{}

### 2018-04-20 article/article.tex

@@
 \usepackage{titlesec}
-% Roman numerals for the sections
-\renewcommand\thesection{\Roman{section}}
-% roman numerals for subsections
-\renewcommand\thesubsection{\roman{subsection}}
-% Change the look of the section titles
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
-% Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{}
 
@@
 \right.
-$$
+\end{equation}
+
+\subsubsection{Objective function}
+
+With a cost defined for partition and transition, it is now possible to
+aggregate everything in order to build a cost function over an entire
+path. For a path ($path = [P_0, ..., P_n]$) with a time from $t_0$ to $t_n$,
+the cost function is given by:
+\begin{equation}
+  \begin{split}
+  f(path) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\ \theta \times C_{tr}(P_{i-1}, P_i)]   
+  \end{split}
+\end{equation}
+
+with $\theta > 0$ a parameter to determine.
+This is the imediate function to minimize in the A$^*$ algorithm.
+
+The Monte Carlo Tree search algorithm maximize a objective function. This function
+represent a potential reward over a path, i.e. a list of partitions over time.
+Given the loss function defined previously, the target function can be
+construct as:
+\begin{equation}
+  \psi(path) = \frac{1}{f(path)}
+\end{equation}
 

### 2018-04-21 article/article.tex
(~) Simplified a lot layout
latex defaults are not that bad
(~) Corrections in intro

@@
-\documentclass[twoside,twocolumn]{article}
+\documentclass[oneside,twocolumn]{article}
 
@@
 
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
-\usepackage{booktabs}
-
-\usepackage{lettrine}
-
-\usepackage{enumitem}
-\setlist[itemize]{noitemsep}
-
-\usepackage{abstract}
-% Set the "Abstract" text to bold
-\renewcommand{\abstractnamefont}{\normalfont\bfseries}
-% Set the abstract itself to small italic text
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
-
-% Allows customization of titles
-\usepackage{titlesec}
-
-% Headers and footers
-\usepackage{fancyhdr}
-% All pages have headers and footers
-\pagestyle{fancy}
-% Blank out the default header
-% \fancyhead{}
-% Blank out the default footer
-\fancyfoot{}
-% Custom header text
-% \fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
-% Custom footer text
-\fancyfoot[RO,LE]{\thepage}
-\usepackage{titling} % Customizing the title section
-
-\usepackage{hyperref} % For hyperlinks in the PDF
+%\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
+%\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
+%\usepackage{booktabs}
+%
+%\usepackage{lettrine}
+%
+%\usepackage{enumitem}
+%\setlist[itemize]{noitemsep}
+%
+%\usepackage{abstract}
+%% Set the "Abstract" text to bold
+%\renewcommand{\abstractnamefont}{\normalfont\bfseries}
+%% Set the abstract itself to small italic text
+%\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
+%
+%% Allows customization of titles
+%\usepackage{titlesec}
+%
+%% Headers and footers
+%\usepackage{fancyhdr}
+%% All pages have headers and footers
+%\pagestyle{fancy}
+%% Blank out the default header
+%% \fancyhead{}
+%% Blank out the default footer
+%\fancyfoot{}
+%% Custom header text
+%% \fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
+%% Custom footer text
+%\fancyfoot[RO,LE]{\thepage}
+%\usepackage{titling} % Customizing the title section
 
@@
 \date{\today}
+\pagestyle{headings}
 
@@
 
-% Move the title up
-\setlength{\droptitle}{-4\baselineskip}
-% Article title formatting
-\pretitle{\begin{center}\Huge\bfseries}
-% Article title closing formatting
-  \posttitle{\end{center}}
+%% Move the title up
+%\setlength{\droptitle}{-4\baselineskip}
+%% Article title formatting
+%\pretitle{\begin{center}\Huge\bfseries}
+%% Article title closing formatting
+%  \posttitle{\end{center}}
 
-\renewcommand{\maketitlehookd}{%
-  \begin{abstract}
-
-  \end{abstract}
-}
+%\renewcommand{\maketitlehookd}{%
+%  \begin{abstract}
+%
+%  \end{abstract}
+%}
 
 %-----------------------------------------------------------------------------
-
 \begin{document}
@@
 
+\begin{abstract}
+  A nice abstract
+\end{abstract}
+
 \section*{Notations}
-\begin{itemize}
+\begin{compactitem}
   \item $m$: elementary module;
@@
   \item \(\first(\ell)\): returns the first element of \(\ell\).
-\end{itemize}
+\end{compactitem}
 
@@
 Currently, configuration is mainly decided on the fly by chief officer (TODO).
-This decision is based on the actual workload on each position. This intuitive
-approach does not take into account future traffic. This papers aims at
-providing a method offering a smoother workload during the day.
+This decision is based on the actual workload on each position. In this
+approach, future workload estimation is based on the controller's feelings and
+it therefore lacks a workload prediction tool.
+This papers aims at providing this tool as well as a method offering a smoother
+workload during the day.
 
@@
 
-As seen in~\cite{treimuth2016branch}
-or~\cite{sergeeva2017dynamic}, a temporal sequence of configurations will be
-provided. Two costs will be considered to create the sequence, namely the cost
-associated to each configuration and transition. The former is based on the
-workload estimated for one sector given a neural network
+In this paper, a temporal sequence of configuration is provided, as done
+in~\cite{treimuth2016branch} or~\cite{sergeeva2017dynamic}.
+Two costs are considered to create the sequence, namely the cost
+associated to each configuration and a transition cost. The former is based on
+the workload estimated for one sector given a neural network
 (see~\cite{gianazza2010forecasting}).
 
-The sequence of configurations will be represented as a tree. The resulting
+The sequence of configurations is represented as a tree. The resulting
 sequence will therefore be a path from the root to a leaf of this tree. This
-problem is highly combinatorial (partitioning of the airspace). Knowing that a
-sochastic tree search algorithm has been honing from a good
-player~\cite{gelly2012grand} to the best one with AlphaGo on one of the most
-highly combinatorial game wich is the game of Go entices us to use one for this
-problem. The Monte Carlo tree search~\cite{browne2012survey} algorithm will be
-used to fullfill this task.
+problem is highly combinatorial (partitioning of the airspace). Since stochastic
+tree search algorithms, combined with deep neural networks have proved
+themselves worthy by outranking the best human player of one of the most highly
+combinatorial game which is the game of Go, the Monte Carlo tree
+search~\cite{browne2012survey} algorithm is used in this paper to fullfill the
+previously mentioned task.
 
-\section{Previous related works}
-\label{sec:previous_related_works}
+\section{Previous related works}\label{sec:previous_related_works}
 
@@
 one\footnote{i.e.\ possibly leading to the best evaluation} according to a
-\emph{tree policy}. This type of problem is addressed by bandits methods where
-a multi armed bandit aims to always activate the slot machine giving the highest
-reward.
+\emph{tree policy}. This type of problem can be solved by bandits methods. Those
+methods consist in, given a bandit in front of several slot machines (multi
+armed bandit), decide which machine will bring the highest reward knowing the
+past result. The objective of bandits methods is thus to maximise the reward and
+minimise the regret of not playing the best machine.
 
-The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
-It uses the UCB (Upper Confidence Bound) as tree policy (introduced
-in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
+The bandit problem has been applied to MCTS via the Upper Confidence Tree
+algorithm in~\cite{kocsis2006bandit} using the UCB1 equation~\ref{eq:ucb1}.
+Let \(v\) be the node from which an action must be
 selected to go deeper in the tree, \(a\) an action available from \(v\),
@@
 \begin{equation}
+  \label{eq:ucb1}
   \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
@@
 While the previous equation is well suited for two players games, it can be
-tweaked to improve its efficience in one player games or sequencing problems. An
+tweaked to improve its efficiency in one player games or sequencing problems. An
 alternative using the standard deviation of the outcome of the simulations is
@@
   \sqrt{%
-    \frac{\beta\log T_v}{t_{v,a}}
+    \frac{\beta\log T_v}{T_{v,a}}
     \min\left( \frac{1}{4}, \sigma^2_{v,a} +

### 2018-04-21 article/article.tex
(~) Some corrections in model part
path -> \pi, array -> cases, no \times sign

@@
 \begin{itemize}
-  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
-  \item $c_=(P, t) = 1 / (\sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$)
-  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
+  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) p_{high}^{S, t} |S|^2$
+  \item $c_=(P, t) = 1 / \left(
+      \sum_{S \in P_t} \delta_{n}(S, t) p_{norm}^{S} |S|^{-2}
+    \right)$
+  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) p_{low}^{S, t} |S|^{-2}$
 \end{itemize}
@@
 regarding $c_+$, $c_=$, $c_-$ and $n$ (cardinal of the partition $P$). The
-partition cost is defined as follow:
+partition cost is defined as follows:
 \begin{equation}
@@
 In a real application, it maybe interesting to order those parmaeters as
-follow : $\alpha > \gamma > \beta > \lambda$.
+follow: $\alpha > \gamma > \beta > \lambda$.
 
@@
   C_{tr}(P_1, P_2) =
-\left\{
-\begin{array}{l}
-    0 \qquad if \qquad P_1 = P_2 \\
-    1 \qquad otherwise
-\end{array}
-\right.
+  \begin{cases}
+    0 & \text{if\,} P_1 = P_2\\
+    1 & \text{otherwise}
+  \end{cases}
 \end{equation}
@@
 aggregate everything in order to build a cost function over an entire
-path. For a path ($path = [P_0, ..., P_n]$) with a time from $t_0$ to $t_n$,
+path. For a path $\pi = [P_0, \dots, P_n]$ with a time from $t_0$ to $t_n$,
 the cost function is given by:
@@
   \begin{split}
-  f(path) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\ \theta \times C_{tr}(P_{i-1}, P_i)]   
+    f(\pi) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\
+    \theta C_{tr}(P_{i-1}, P_i)]
   \end{split}
@@
 with $\theta > 0$ a parameter to determine.
-This is the imediate function to minimize in the A$^*$ algorithm.
+This is the immediate function to minimize in the A$^*$ algorithm.
 
 The Monte Carlo Tree search algorithm maximize a objective function. This function
-represent a potential reward over a path, i.e. a list of partitions over time.
+represent a potential reward over a path, i.e.\ a list of partitions over time.
 Given the loss function defined previously, the target function can be
@@
 \begin{equation}
-  \psi(path) = \frac{1}{f(path)}
+  \psi(\pi) = \frac{1}{f(\pi)}
 \end{equation}

### 2018-04-23 article/article.tex
(~) Refactored notations
No action, only nodes

@@
   \item $C(P, t)$: cost of a given partition $P$ at time $t$;
-  \item \(\mathcal{S}\): the set of available states;
-  \item \(\mathcal{A}\): the set of all possible actions;
-  \item \(\mathcal{A}(s)\): the available actions from a state \(s\);
-  \item $s$: state;
-  \item \(f \colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\):
-    transition function giving the resulting states of action \(a\) executed
-    from a state \(s\).
-  \item $h \colon \mathcal{S} \to \mathbb{R}^+$: heuristic estimating the cost
-    from a state \(s \in \mathcal{S}\) to a final state;
-  \item \(\mathcal{N}\): the set of nodes
   \item \(u, v \in \mathcal{N}^2\): nodes
-  \item \(Q \colon \mathcal{N} \to \mathbb{R}^+\): the reward expectancy of a
-    node;
-  \item \(N \colon \mathcal{N} \to \mathbb{N}\): number of simulations running
-    through or from a node;
+  \item \(\mathcal{N}\): the set of nodes;
+  \item $h \colon \mathcal{N} \to \mathbb{R}^+$: heuristic estimating the cost
+    from a node \(u\) to a final node;
+  \item \(\mu_v\): mean value of outcomes of simulations run through or from a
+    node \(v\);
+  \item \(T_v\): number of simulations run through or from a node \(v\);
   \item \(\first(\ell)\): returns the first element of \(\ell\).
@@
 \item expansion: adding new nodes to the search tree,
-\item simulation (or random walk): choosing quickly actions from the model
+\item simulation (or random walk): choosing successively nodes from the model
   tree,
@@
 \paragraph{Selection (algorithm~\ref{alg:tree_pol})}
-The aim of the selection is to choose from a set of actions, the most promising
-one\footnote{i.e.\ possibly leading to the best evaluation} according to a
-\emph{tree policy}. This type of problem can be solved by bandits methods. Those
-methods consist in, given a bandit in front of several slot machines (multi
-armed bandit), decide which machine will bring the highest reward knowing the
-past result. The objective of bandits methods is thus to maximise the reward and
-minimise the regret of not playing the best machine.
+The aim of the selection is to choose the most promising node among the children
+available\footnote{i.e.\ possibly leading to the best evaluation} according to
+a \emph{tree policy}. This type of problem can be solved by bandits methods.
+Those methods consist in, given a bandit in front of several slot machines
+(multi armed bandit), decide which machine will bring the highest reward knowing
+the past result. The objective of bandits methods is thus to maximise the reward
+and minimise the regret of not playing the best machine.
 
@@
 algorithm in~\cite{kocsis2006bandit} using the UCB1 equation~\ref{eq:ucb1}.
-Let \(v\) be the node from which an action must be
-selected to go deeper in the tree, \(a\) an action available from \(v\),
-\(\mu_{v,a}\) the expected outcome from the node obtained by applying action
-\(a\) on \(v\), \(T_v\) the number of simulations carried out from node \(v\)
-(which includes any simulation from nodes in any subtree of \(v\)) and
-\(T_{v,a}\) the number of simulations carried out from the node resulting from
-the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
-is thus the one maximising the \emph{UCB1} equation
+Let \(u\) be the node from which a child \(v\) must be selected to go deeper in
+the tree, \(T_u\) the number of simulations carried out from node \(u\)
+(which includes any simulation from nodes in any subtree of \(u\)) and
+\(\beta\) a constant. The selected child \(v\) is thus the one maximising the
+\emph{UCB1} equation
 \begin{equation}
   \label{eq:ucb1}
-  \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
+  \mu_v + 2 \beta \sqrt{\frac{2 \log T_u}{T_v}}
 \end{equation}
@@
 \begin{equation}
-  \mu_{v,a} +
+  \mu_v +
   \sqrt{%
-    \frac{\beta\log T_v}{T_{v,a}}
-    \min\left( \frac{1}{4}, \sigma^2_{v,a} +
+    \frac{\beta\log T_u}{T_v}
+    \min\left( \frac{1}{4}, \sigma^2_v +
     \sqrt{%
-      \frac{2 \log T_v}{T_{v,a}}
+      \frac{2 \log T_u}{T_v}
     }\right)
@@
     \Function{Expand}{$v$}
-    \State{}$\ell \gets \{ v | v \text{\,children of\,} u, N(v) = 0 \}$
+    \State{}$\ell \gets \{ v | v \text{\,children of\,} u, T_v = 0 \}$
     \State{}\Return{random element from $\ell$}
@@
   \begin{algorithmic}
-    \Procedure{Backpropagate}{$v$, $r$}
-    \State{}\(\delta \gets\ r - \mu_v\)
-    \State{}\(Q(v) \gets Q(v) + \frac{\delta}{N(v) + 1}\)
-    \State{}\(N(v) \gets N(v) + 1\)
+    \Procedure{Backpropagate}{$u$, $r$}
+    \State{}\(\delta \gets r - \mu_u\)
+    \State{}\(\mu_u \gets \mu_u + \frac{\delta}{T_u + 1}\)
+    \State{}\(T_u \gets T_u + 1\)
     \EndProcedure{}
@@
 For the UCB1-tuned, the standard deviation has to be computed. It results in
-algorithm~\ref{alg:backpropagate-tuned} which introduces the value \(m_2(v)\)
-associated to a node \(v\) to compute the standard deviation via the formula
+algorithm~\ref{alg:backpropagate-tuned} which introduces the value \(m_2(u)\)
+associated to a node \(u\) to compute the standard deviation via the formula
 \begin{equation}
-  \sigma^2_v = \frac{m_2(v)}{N(v)}.
+  \sigma^2_u = \frac{m_2(u)}{T_u}.
 \end{equation}
@@
   \begin{algorithmic}
-    \Procedure{Backpropagate}{$v, r$}
-    \State{}\(\delta \gets r - Q(v)\)
-    \State{}\(Q(v) \gets Q(v) + \frac{\delta}{N(v) + 1}\)
-    \State{}\(\delta' \gets r - Q(v)\)
-    \State{}\(m_2(v) \gets m_2(v) + \delta \delta'\)
-    \State{}\(N(v) \gets N(v) + 1\)
+    \Procedure{Backpropagate}{$u, r$}
+    \State{}\(\delta \gets r - \mu_u\)
+    \State{}\(\mu_u \gets \mu_u + \frac{\delta}{T_u + 1}\)
+    \State{}\(\delta' \gets r - \mu_u\)
+    \State{}\(m_2(u) \gets m_2(u) + \delta \delta'\)
+    \State{}\(T_u \gets T_u + 1\)
     \EndProcedure{}
@@
 \subsubsection{Node selection}\label{sssec:node_selection}
-The building of the final sequence of actions differs from the selection on the
-aim, the selection chooses the most promising actions while the final sequence
+The building of the final sequence of nodes differs from the selection on the
+aim, the selection chooses the most promising node while the final sequence
 must be composed of the best. This means that the final path build won't use the
@@
     \Function{NextNode}{$u$}
-    \State{}\Return{$\argmax\{Q(v) | v \text{\,children of\,} u\}$}
+    \State{}\Return{$\argmax\{\mu_v | v \text{\,children of\,} u\}$}
     \EndFunction{}

### 2018-04-23 article/article.tex
(+) Heuristic A* some corrections

@@
     \While{within computational budget}
-    \State{} \([v_0, \dots, v_n] \gets\) \Call{TreePolicy}{$v_0$}
-    \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v_n$}
-    \State{} \Call{Backup}{$[v_0, \dots, v_n]$, $\Delta$}
+    \State{}\(\pi \gets\) \Call{TreePolicy}{$v_0$}
+    \State{}\(v \gets \first(\pi)\)
+    \State{}\(\Delta \gets\) \Call{DefaultPolicy}{$v$}
+    \State{}\Call{Backup}{$\pi$, $\Delta$}
     \EndWhile{}
@@
 \end{algorithm}
+
+\subparagraph{Exploration exploitation tradeoff} The constant \(\beta\) answers
+to the exploration v exploitation dilemma. In the UCB1 equation~\ref{eq:ucb1},
+the left hand term increases the value of less explorated nodes to consider them
+as still promising. A high \(\beta\) value will therefore make the algorithm
+prone to try unvisited nodes while a low value will consider almost exclusively
+the results obtained so far, even if other paths are better.
 
@@
 nodes, \(h\) a heuristic function estimating the cost from a state \(u\) given
-as argument to a final state, \(k \colon \mathcal{S}^2 \to \mathbb{R}\). The
-function \(\mathup{first(G)}\) returns the first element of \(G\) and
+as argument to a final state, \(k \colon \mathcal{N}^2 \to \mathbb{R}\). The
+function \(\first(G)\) returns the first element of \(G\) and
 \(f\mathup{-insert}(G, v)\) inserts \(v\) in \(G\) ordering first by \(f\)
@@
     \While{$G \neq \emptyset$}
-    \State{}\(u \gets \mathup{first}(G); G \gets G \backslash \{u\}\)
+    \State{}\(u \gets \first(G); G \gets G \backslash \{u\}\)
     \State{}\(D \gets D \cup \{u\}\)
@@
 \end{algorithm}
+
+\paragraph{Heuristic}
+To be sure to have the optimal solution, the heuristic \(h\) has to be
+\emph{minimal} i.e.\ let \(h^*\) be the optimal heuristic (the one giving the
+true distance from a node to a final state), then for any node \(u\), \(h(u)
+\leq h^*(u)\).
+
 \section{Model}

### 2018-04-23 article/article.tex
(+) partitions + transitions

--- 2018-04-21 article/article.tex
@@
 
-\subsection{workload}
+\subsubsection{Partitions}
+
+In the configuration problem, the airspace needs to be devided into sectors. Each
+sector represent a area controlled by two officers and is composed of
+one or more elementary modules. A partition $P$ is a set of
+sectors covering all the airspace.
+
+\subsubsection{Transitions}
+
+During the day, and with the evolution of the trafic, the workload (rigorously
+defined in \ref{sub:workload}) vary. To help controllers to maintain a
+homogeneous level of workload, the airspace partitioning needs to be adapt
+and therefore make transition from an inital partition to a target partition.
+This maneuver requires coordination hence increase controller's workload.
+This is the reason why a full reconfiguration isn't doable in practice, and
+only a selection of transitions can be operate.
+
+In our model, three transistions are possible namely a merge, a transfer
+and a split. For an airspace composed of three modules $A$, $B$, and $C$,
+those actions represent:
+\begin{itemize}
+  \item merge: $\{\{A, B\}, \{C\}\} \quad \rightarrow \quad \{\{A, B, C\}\}$ 
+  \item split: $\{\{A, B, C\}\} \quad \rightarrow \quad \{\{A, B\}, \{C\}\}$ 
+  \item transfer: $\{\{A, B\}, \{C\}\} \quad \rightarrow \quad \{\{A\}, \{B, C\}\}$ 
+\end{itemize}
+
+\subsection{Workload}\label{sub:workload}
 
@@
 The Monte Carlo Tree search algorithm maximize a objective function. This function
-represent a potential reward over a path, i.e.\ a list of partitions over time.
+represent the reward of a path, i.e.\ a list of partitions over time.
 Given the loss function defined previously, the target function can be
@@
 \begin{equation}
-  \psi(\pi) = \frac{1}{f(\pi)}
+  Q(\pi) = \frac{1}{f(\pi)}
 \end{equation}

### 2018-04-23 article/article.tex
(+) todo
(+) par on heuristic

--- 2018-04-23 article/article.tex
@@
 
+\subsection{Heuristic}
+To use effectively the A\(^*\) algorithm, a heuristic is needed. Using the null
+heuristic (\(\forall u \in \mathcal{N}, h(u) = 0\)) revealed to be inefficient
+considering the branching factor. To build a minimal heuristic, the limitations
+introduced in~\ref{ssec:partitions} are not considered, neither those associated
+to the previous partition (split, merge or transfer). This allows to generate
+the sequence of the best partitions among all possible at each time step.
+
 \bibliography{article}

### 2018-04-23 article/article.tex

--- 2018-04-23 article/article.tex
@@
 
+TODO : parler de la restriction des groupements de modules
+
 \subsubsection{Transitions}

### 2018-04-26 article/article.tex
(+) added ref to partitions par
(+) Personal settings regarding headings
(+) added par on theoretical performances

@@
 \usepackage{paralist}
+\usepackage{titlesec}
 
@@
 
-%\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
-%\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
-%\usepackage{booktabs}
-%
-%\usepackage{lettrine}
-%
-%\usepackage{enumitem}
-%\setlist[itemize]{noitemsep}
-%
-%\usepackage{abstract}
-%% Set the "Abstract" text to bold
-%\renewcommand{\abstractnamefont}{\normalfont\bfseries}
-%% Set the abstract itself to small italic text
-%\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
-%
-%% Allows customization of titles
-%\usepackage{titlesec}
-%
-%% Headers and footers
-%\usepackage{fancyhdr}
-%% All pages have headers and footers
-%\pagestyle{fancy}
-%% Blank out the default header
-%% \fancyhead{}
-%% Blank out the default footer
-%\fancyfoot{}
-%% Custom header text
-%% \fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1}
-%% Custom footer text
-%\fancyfoot[RO,LE]{\thepage}
-%\usepackage{titling} % Customizing the title section
+% Headings customisation
+\titleformat*{\section}{\Large\scshape\center}
+\renewcommand{\thesection}{\Roman{section}}
+\renewcommand{\thesubsection}{\Alph{subsection}}
 
@@
 
-%-----------------------------------------------------------------------------
-%	TITLE SECTION
-%-----------------------------------------------------------------------------
-
-%% Move the title up
-%\setlength{\droptitle}{-4\baselineskip}
-%% Article title formatting
-%\pretitle{\begin{center}\Huge\bfseries}
-%% Article title closing formatting
-%  \posttitle{\end{center}}
-
-%\renewcommand{\maketitlehookd}{%
-%  \begin{abstract}
-%
-%  \end{abstract}
-%}
-
-%-----------------------------------------------------------------------------
 \begin{document}
@@
 
+\subsection{Theoretical performances}
+In this paper, we consider a high branching factor problem.
+
+Let \(M\) be the number of edges between the inital state and a final state and
+\(K\) the branching factor.
+Then, the complexity of the A\(^*\) is expected to be, in the worst case,
+exponential in \(M\) (see~\cite{alliotschiex2002ia&it}).
+
+Since the Monte carlo tree is a stochastic method, one must wait for a
+satisfactorily converging solution rather than the exact solution. It has
+however been proven in~\cite{kocsis2006bandit} that the bias of the expected
+payoff is \(O\left(\frac{\log n}{n}\right)\). Furthermore, the failure
+probability\footnote{probability of not selecting the best child} converges to
+zero at a polynomial rate.
+
 \section{Model}
@@
 
-\subsubsection{Partitions}
+\subsubsection{Partitions}\label{ssec:partitions}
 

### 2018-04-26 article/article.tex
(+) workload paragraph

--- 2018-04-23 article/article.tex
@@
 
-TODO : parler de la restriction des groupements de modules
+A sector is a group of elementary modules, but in practice only a collection
+of sectors are allowed in this model. This collection is referred as the context.
+This restriction reflect the fact the shape of a sector needs have a relatively
+'nice' shape in order to be easely manageable for a controller.
 
@@
 
+TODO : meilleur explication des transition
+
 \subsection{Workload}\label{sub:workload}
 
-$p_{low}$
-$p_{normal}$
-$p_{high}$
+Let us now precise the model established to evaluate a workload metric. The
+workload depends on external criteria such as sector and trafic complexity
+or humain factors (e.g. controller fatigue, stress). In our model, we assum
+that the workload felt by a controller is directly and only related to the trafic
+complexity.
+
+The simplest approch considers only the number of aircraft per sector $Nb_{aircraft}$. We
+defined two arbitary thresholds, namely $th_{low}$ and $th_{high}$, defining
+three zones where the workload is low (resp. normal and high) if $Nb_{aircraft} < th_{low}$
+(resp. $th_{low} < Nb_{aircraft} < th_{high}$ and $th_{high} < Nb_{aircraft}$).
+Those workload levels are translated to probabilities $p_{low}$, $p_{normal}$ and
+$p_{high}$, where $p_{low}(S)$ is the probability that the sector $S$ is underloaded
+(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $Nb_{aircraft}$:
+\begin{equation}
+  \begin{cases}
+    p_{low}^{S, t} = \mathbb{1}_{[0, th_{low}]}(Nb_{aircraft}) \\
+    p_{normal}^{S, t} = \mathbb{1}_{[th_{low}, th_{high}]}(Nb_{aircraft})\\
+    p_{high}^{S, t} = \mathbb{1}_{[th_{low}, \infty[}(Nb_{aircraft})
+  \end{cases}
+\end{equation}
+where $\mathbb{1}$ is the indicator function defined by:
+\begin{equation}
+  \mathbb{1}_{I}(x) = 
+  \begin{cases}
+    1 & \text{if\,} \quad x \in I\\
+    0 & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+This expression of the workload, expressed in term of probabilities, allow us to
+use a more complex model for the workload such as a neural network
+(see~\cite{gianazza2010forecasting}). This method learn the probabilities $p_{low}$,
+$p_{normal}$ and $p_{high}$ based on parameters including trafic complexity and
+sector complexity (e.g. airspace volume).
 
@@
   \begin{cases}
-    0 & \text{if\,} P_1 = P_2\\
+    0 & \text{if\,} \quad P_1 = P_2\\
     1 & \text{otherwise}

### 2018-04-26 article/article.tex
(+) transition sentence exposing our model

@@
 (in~\cite{bedouet2016towards}).
+
+In this paper, each partition scheme is based upon the number of aircraft in
+each
+elementary module and the cost of transition from the previous partitioning. The
+available partitions come from a predetermined set containing feasible
+partitionings.
 

### 2018-04-26 article/article.tex
(+) results

@@
 (in~\cite{bedouet2016towards}).
-
-In this paper, each partition scheme is based upon the number of aircraft in
-each
-elementary module and the cost of transition from the previous partitioning. The
-available partitions come from a predetermined set containing feasible
-partitionings.
 
@@
 
-\subsection{Theoretical performances}
-In this paper, we consider a high branching factor problem.
-
-Let \(M\) be the number of edges between the inital state and a final state and
-\(K\) the branching factor.
-Then, the complexity of the A\(^*\) is expected to be, in the worst case,
-exponential in \(M\) (see~\cite{alliotschiex2002ia&it}).
-
-Since the Monte carlo tree is a stochastic method, one must wait for a
-satisfactorily converging solution rather than the exact solution. It has
-however been proven in~\cite{kocsis2006bandit} that the bias of the expected
-payoff is \(O\left(\frac{\log n}{n}\right)\). Furthermore, the failure
-probability\footnote{probability of not selecting the best child} converges to
-zero at a polynomial rate.
-
 \section{Model}
@@
 
+\section{Results}
+
+TODO
+
 \bibliography{article}

### 2018-04-26 article/article.tex
(-) spelling mistakes, (~) title

--- 2018-04-26 article/article.tex
@@
 
-\title{Airspace configuration decision based on Monte-Carlo tree search}
+\title{Monte Carlo tree search compared to A\(^*\) in airspace configuration
+decision problem}
 \author{Gabriel~Hondet, Beno\^{\i}t Viry}
@@
 combinatorial game which is the game of Go, the Monte Carlo tree
-search~\cite{browne2012survey} algorithm is used in this paper to fullfill the
+search~\cite{browne2012survey} algorithm is used in this paper to fulfill the
 previously mentioned task.
@@
 available partitions come from a predetermined set containing feasible
-partitionings.
+partitioning.
 
@@
 \end{compactenum}
-Those phases are repeated until a stopping criterium (e.g.\ memory or time) is
+Those phases are repeated until a stopping criterion (e.g.\ memory or time) is
 reached, resulting in algorithm~\ref{alg:gen_mcts} where the tree policy
@@
 
-\subparagraph{Exploration exploitation tradeoff} The constant \(\beta\) answers
+\subparagraph{Exploration exploitation trade-off} The constant \(\beta\) answers
 to the exploration v exploitation dilemma. In the UCB1 equation~\ref{eq:ucb1},
-the left hand term increases the value of less explorated nodes to consider them
+the left hand term increases the value of less explored nodes to consider them
 as still promising. A high \(\beta\) value will therefore make the algorithm

### 2018-05-01 article/article.tex
(+) Completed first draft

@@
   \item $C(P, t)$: cost of a given partition $P$ at time $t$;
+  \item \(C_{tr}(P_1, P_2)\): cost of transition between partitions \(P_1\) and
+    \(P_2\);
+  \item \(\pi = [P_1, \dots, P_n]\): a sequence of partitions (called \emph{path}
+      in the tree search algorithms);
+  \item \(f(\pi)\): cost associated to sequence of partitions \(\pi\);
+  \item \(\mathcal{N}\): the set of nodes;
   \item \(u, v \in \mathcal{N}^2\): nodes
-  \item \(\mathcal{N}\): the set of nodes;
   \item $h \colon \mathcal{N} \to \mathbb{R}^+$: heuristic estimating the cost
@@
 \subsection{Monte Carlo tree search}
-\subsubsection{Overview}
+\subsubsection{Overview}\label{sssec:overview}
 Monte-Carlo is a best-first search method using stochastic simulations. The
+algorithm is based on the computation of the reward expectancy of paths which
+is estimated through Monte-Carlo simulations.
+The
 algorithm actually uses two trees, an underlying tree associated to the model
-(e.g.\ a game tree) and a search tree. The latter is built incrementally by the
-algorithm the following way, at each step, starting from the root, the most
-promising nodes towards the bottom of the tree are selected. Once a leaf of the
-search tree is reached, a new node coming from the model tree is added to the
-leaf. A simulation is then run from this newly added node to evaluate the
-outcome of a path starting from it. The result of the simulation is eventually
-backpropagated to the consecutive parents of the new node.
-
-The above mentioned steps are denoted
+(e.g.\ a game tree) and a search tree. The latter is built incrementally, each
+step being composed of four phases, namely
 \begin{compactenum}
-\item selection (or tree walk): choosing successively most promising nodes,
-\item expansion: adding new nodes to the search tree,
+\item\label{it:selection} selection (or tree walk): choosing successively
+  most promising nodes from the search tree,
+\item\label{it:expansion} expansion: adding new nodes to the search tree,
 \item simulation (or random walk): choosing successively nodes from the model
-  tree,
+  tree, from the expanded node to a leaf,
 \item backpropagation (or backup): applying the result of the simulation to
-  the search tree.
+  the previously selected nodes of the search tree (phases~\ref{it:selection}
+  and~\ref{it:expansion}).
 \end{compactenum}
@@
 \paragraph{Selection (algorithm~\ref{alg:tree_pol})}
-The aim of the selection is to choose the most promising node among the children
-available\footnote{i.e.\ possibly leading to the best evaluation} according to
-a \emph{tree policy}. This type of problem can be solved by bandits methods.
+The aim of the selection is to build a path from the root of the search tree
+by choosing successively the most
+promising\footnote{i.e.\ possibly leading to the best evaluation} node.
+Given a node \(u\) that has previously been selected, the
+best node --
+according to a \emph{tree policy} -- among the children of \(u\) is
+chosen.
+This type
+of problem can be solved by bandits methods.
 Those methods consist in, given a bandit in front of several slot machines
-(multi armed bandit), decide which machine will bring the highest reward knowing
-the past result. The objective of bandits methods is thus to maximise the reward
+(multi armed bandit), deciding which machine will bring the highest reward
+knowing
+the past results. The objective of bandits methods is thus to maximise the reward
 and minimise the regret of not playing the best machine.
@@
 (which includes any simulation from nodes in any subtree of \(u\)) and
-\(\beta\) a constant. The selected child \(v\) is thus the one maximising the
+\(\beta\) a chosen constant. The selected child \(v\) is the one maximising the
 \emph{UCB1} equation
@@
 tweaked to improve its efficiency in one player games or sequencing problems. An
-alternative using the standard deviation of the outcome of the simulations is
+alternative using the standard deviation \(\sigma_v\) of the outcome of the
+previous simulations involving node \(v\) is
 proposed in~\cite{sebag2010fuse} called the \emph{UCB1-tuned} equation
-\begin{equation}
+\begin{equation}\label{eq:ucb1-tuned}
   \mu_v +
-  \sqrt{%
-    \frac{\beta\log T_u}{T_v}
+  \beta\sqrt{%
+    \frac{\log T_u}{T_v}
     \min\left( \frac{1}{4}, \sigma^2_v +
@@
 \subparagraph{Exploration exploitation trade-off} The constant \(\beta\) answers
-to the exploration v exploitation dilemma. In the UCB1 equation~\ref{eq:ucb1},
-the left hand term increases the value of less explored nodes to consider them
+to the exploration-exploitation dilemma. In the UCB1 equations~\ref{eq:ucb1}
+and~\ref{eq:ucb1-tuned},
+the right hand term increases the UCB value of less explored nodes to consider them
 as still promising. A high \(\beta\) value will therefore make the algorithm
 prone to try unvisited nodes while a low value will consider almost exclusively
-the results obtained so far, even if other paths are better.
+the results obtained so far, even if other paths are better but unexplored.
 
 \paragraph{Expansion (algorithm~\ref{alg:expansion})}
-The above step is carried out only if, considering a node \(v\), a simulation
-has been run through each child available from \(v\). In the case where one or
-more children haven't been explored yet, one of the unexplored children is
-chosen and a simulation is started from it.
+If the selected node has one or more unvisited children, the selection stops and
+one of them is added to the search tree randomly, the latter being thus expanded
+by this new node.
 \begin{algorithm}
@@
 \subsubsection{Simulation and backpropagation}
+The two steps described in this paragraph aim to guess the reward that can be
+expected from a path including a given node.
 \paragraph{Simulation}
@@
 \subsection{Sequence building}
-\subsubsection{Node selection}\label{sssec:node_selection}
-The building of the final sequence of nodes differs from the selection on the
-aim, the selection chooses the most promising node while the final sequence
-must be composed of the best. This means that the final path build won't use the
-UCB equation. Chaslot \textit{et al.}~\cite{browne2012survey} propose several
+Once the stopping criterium mentioned in~\ref{sssec:overview} is matched, the
+sequence of states can be extracted from the search tree.
+
+\subsubsection{Choosing nodes}\label{sssec:node_selection}
+To build the final sequence, nodes are chosen according to a criterion.
+Chaslot \textit{et al.} in~\cite{chaslot2008progstrat} propose several
 methods to select nodes,
 \begin{compactitem}
-  \item max-child: select the child with highest reward;
+  \item max-child: select the child with highest mean reward;
   \item robust child: select the most visited root child;
@@
 matched, a node \(v\) among the ones reachable from \(u\) is selected via one of
-the previously mentioned policies. Then an mcts is called back with node \(u\)
+the previously mentioned policies. Then the MCTS algorithm is called back with
+node \(u\)
 as the root node. The path is composed of the successive roots. The algorithm is
@@
 \subsection{Theoretical performances}
-In this paper, we consider a high branching factor problem.
+In this paper, a high branching factor problem is considered.
 
-Let \(M\) be the number of edges between the inital state and a final state and
-\(K\) the branching factor.
+Let \(H\) be the height of the tree, \(K\) the branching factor.
 Then, the complexity of the A\(^*\) is expected to be, in the worst case,
-exponential in \(M\) (see~\cite{alliotschiex2002ia&it}).
+exponential in \(H\) (see~\cite{alliotschiex2002ia&it}).
 
@@
 however been proven in~\cite{kocsis2006bandit} that the bias of the expected
-payoff is \(O\left(\frac{\log n}{n}\right)\). Furthermore, the failure
-probability\footnote{probability of not selecting the best child} converges to
-zero at a polynomial rate.
+payoff tends to zero.
 
@@
 
+\subsubsection{Time step}
+Since the overall goal is to compute a temporal sequence of partitions over a
+day, the day has to be sampled. This sampling in represented as the temporal
+difference between a node and its children in the trees mentioned in the
+previous section.
+
 \subsubsection{Partitions}\label{ssec:partitions}
 
-In the configuration problem, the airspace needs to be devided into sectors. Each
-sector represent a area controlled by two officers and is composed of
+In the configuration problem, the airspace needs to be divided into sectors. Each
+sector represents an area controlled by two controllers and is composed of
 one or more elementary modules. A partition $P$ is a set of
@@
 of sectors are allowed in this model. This collection is referred as the context.
-This restriction reflect the fact the shape of a sector needs have a relatively
-'nice' shape in order to be easely manageable for a controller.
+This restriction reflects the fact the more complex the shape is, the harder it
+is to manage.
 
-\subsubsection{Transitions}
+\subsubsection{Transitions}\label{sssec:transitions}
 
 During the day, and with the evolution of the trafic, the workload (rigorously
-defined in \ref{sub:workload}) vary. To help controllers to maintain a
-homogeneous level of workload, the airspace partitioning needs to be adapt
-and therefore make transition from an inital partition to a target partition.
-This maneuver requires coordination hence increase controller's workload.
-This is the reason why a full reconfiguration isn't doable in practice, and
-only a selection of transitions can be operate.
+defined in~\ref{sub:workload}) vary. To help controllers to maintain a
+homogeneous level of workload, the airspace partitioning needs to be adapted
+through a transition from an inital partition to a target partition.
+This maneuver requires either coordination or the opening of a new position
+and hence increases controller's workload.
+This is the reason why a full reconfiguration isn't feasible, and
+only a subset of all available transitions can be operated.
 
-In our model, three transistions are possible namely a merge, a transfer
-and a split. For an airspace composed of three modules $A$, $B$, and $C$,
+In the model considered here, three transistions are possible namely a merge,
+a transfer
+or a split. For an airspace composed of three modules $A$, $B$, and $C$,
 those actions represent:
@@
 
-TODO : meilleur explication des transition
+TODO : meilleure explication des transitions?
+
+Each transition is composed of at most one merge, split or tranfer; more would
+result in too complex reconfigurations.
 
@@
 workload depends on external criteria such as sector and trafic complexity
-or humain factors (e.g. controller fatigue, stress). In our model, we assum
-that the workload felt by a controller is directly and only related to the trafic
+or humain factors (e.g. controller health or stress). In the model assumed here,
+the workload felt by a controller is directly and only related to the trafic
 complexity.
 
-The simplest approch considers only the number of aircraft per sector $Nb_{aircraft}$. We
+The simplest approch considers only the number of aircraft per sector $N_{aircraft}$. We
 defined two arbitary thresholds, namely $th_{low}$ and $th_{high}$, defining
-three zones where the workload is low (resp. normal and high) if $Nb_{aircraft} < th_{low}$
-(resp. $th_{low} < Nb_{aircraft} < th_{high}$ and $th_{high} < Nb_{aircraft}$).
+three zones where the workload is low (resp. normal and high) if $N_{aircraft} < th_{low}$
+(resp. $th_{low} < N_{aircraft} < th_{high}$ and $th_{high} < N_{aircraft}$).
 Those workload levels are translated to probabilities $p_{low}$, $p_{normal}$ and
 $p_{high}$, where $p_{low}(S)$ is the probability that the sector $S$ is underloaded
-(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $Nb_{aircraft}$:
+(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_{aircraft}$:
 \begin{equation}
-  \begin{cases}
-    p_{low}^{S, t} = \mathbb{1}_{[0, th_{low}]}(Nb_{aircraft}) \\
-    p_{normal}^{S, t} = \mathbb{1}_{[th_{low}, th_{high}]}(Nb_{aircraft})\\
-    p_{high}^{S, t} = \mathbb{1}_{[th_{low}, \infty[}(Nb_{aircraft})
-  \end{cases}
+  \begin{aligned}
+    p_{low}^{S, t} &= \mathbb{1}_{[0, th_{low}]}(N_{aircraft}) \\
+    p_{normal}^{S, t} &= \mathbb{1}_{[th_{low}, th_{high}]}(N_{aircraft})\\
+    p_{high}^{S, t} &= \mathbb{1}_{[th_{low}, \infty[}(N_{aircraft})
+  \end{aligned}
 \end{equation}
@@
 use a more complex model for the workload such as a neural network
-(see~\cite{gianazza2010forecasting}). This method learn the probabilities $p_{low}$,
+(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_{low}$,
 $p_{normal}$ and $p_{high}$ based on parameters including trafic complexity and
-sector complexity (e.g. airspace volume).
+the complexity of the sector (e.g. airspace volume).
 
@@
 In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
-to be defined. This cost depend on the workload in each the sector. We give the
-following definitions~\cite{ferrari2017} to represent a high
-(respectively normal and low) cost for the partition:
-\begin{itemize}
-  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) p_{high}^{S, t} |S|^2$
-  \item $c_=(P, t) = 1 / \left(
-      \sum_{S \in P_t} \delta_{n}(S, t) p_{norm}^{S} |S|^{-2}
-    \right)$
-  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) p_{low}^{S, t} |S|^{-2}$
-\end{itemize}
-with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
+to be defined. In this paper, this cost depends on the workload in each the
+sector. The definitions given in~\cite{ferrari2017} are used to represent a high
+(respectively normal and low) cost for partition \(P\):
+\begin{equation}
+  \begin{aligned}
+    c_+(P, t) &= \sum_{S \in P_t} \delta_{h}(S, t) \cdot p_{high}^{S, t} \cdot |S|^2\\
+    c_=(P, t) &= \left(
+    \sum_{S \in P_t} \delta_{n}(S, t) \cdot p_{norm}^{S} \cdot |S|^{-2}
+    \right)^{-1}\\
+    c_-(P, t) &= \sum_{S \in P_t} \delta_{l}(S, t) \cdot p_{low}^{S, t} \cdot |S|^{-2}
+  \end{aligned}
+\end{equation}
+with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equals 1
 if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
@@
 value $\alpha$ help to
-minimize $c_+$, hence the overall number of sectors with too much tafic.
-In a real application, it maybe interesting to order those parmaeters as
+minimize $c_+$, hence the overall number of sectors with too much trafic.
+In a real application, it maybe interesting to order those parameters as
 follow: $\alpha > \gamma > \beta > \lambda$.
@@
 
-In an operational context, each reconfiguration raise the workload for
-controllers. This trend needs to be considered and a transition cost needs to
-be defined for this matter. A transition is (TODO), and the transition cost is
-then:
+In an operational context, each reconfiguration increases the workload for
+controllers. This needs to be considered via the definition of a transition
+cost.
+A transition is the difference of two partitions separated by only on time step.
+The transition cost is then
 \begin{equation}
@@
   \begin{cases}
-    0 & \text{if\,} \quad P_1 = P_2\\
+    0 & \text{if} \quad P_1 = P_2\\
     1 & \text{otherwise}
@@
 
-With a cost defined for partition and transition, it is now possible to
+Having defined a cost for partitions and transitions, it is now possible to
 aggregate everything in order to build a cost function over an entire
@@
     f(\pi) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\
-    \theta C_{tr}(P_{i-1}, P_i)]
+    \theta \cdot C_{tr}(P_{i-1}, P_i)]
   \end{split}
@@
 
-The Monte Carlo Tree search algorithm maximize a objective function. This function
-represent the reward of a path, i.e.\ a list of partitions over time.
+The Monte Carlo Tree search algorithm maximizes an objective function. This function
+can be interpreted as the reward of a path.
 Given the loss function defined previously, the target function can be
-construct as:
+constructed as:
 \begin{equation}
@@
 considering the branching factor. To build a minimal heuristic, the limitations
-introduced in~\ref{ssec:partitions} are not considered, neither those associated
+introduced in~\ref{sssec:transitions} are not considered, neither those associated
 to the previous partition (split, merge or transfer). This allows to generate
 the sequence of the best partitions among all possible at each time step.
+
+\section{Results}
+
+TODO
 

### 2018-05-03 article/article.tex
(+) Added state of the art on monte carlo
And low, high, normal and aircraft or \text-ed

@@
 partitioning.
+
+The Monte Carlo tree search algorithm has been quickly used in two players game.
+Only three years after its apparition in 1990, Br\"ugmann applies it to the Go
+game in~\cite{brugmann1993mcgo}. While the Monte-Carlo tree search is still
+extensively used in two players games (and especially the Go game), its
+adaptation to one player game has been worked on. Auer \textit{et al.} propose
+an
+upper confidence bound formula which appears to be more efficient for one player
+games as seen in~\cite{sebag2010fuse}. The introduced formula uses the standard
+deviation of the outcome of the simulations.
+The latter article also uses the rapid
+action value estimation (RAVE) technique to quicken the convergence of the
+algorithm. RAVE uses the ``all moves as first'' heuristic, in which all
+moves\footnote{a move is informally considered as a decision taken regarding
+which state to choose while descending the search tree}
+seen during simulations are considered as a first move. This allows the
+algorithm to update more statistics in one simulation. Gelly \textit{et al.} use
+in~\cite{gelly2012grand} the RAVE technique coupled with several heuristics. The
+heuristics bias the initialisaton of a node in the search tree, pre filling its
+statistics using prior knowledge of the problem.
 
@@
 
-The simplest approch considers only the number of aircraft per sector $N_{aircraft}$. We
-defined two arbitary thresholds, namely $th_{low}$ and $th_{high}$, defining
-three zones where the workload is low (resp. normal and high) if $N_{aircraft} < th_{low}$
-(resp. $th_{low} < N_{aircraft} < th_{high}$ and $th_{high} < N_{aircraft}$).
-Those workload levels are translated to probabilities $p_{low}$, $p_{normal}$ and
-$p_{high}$, where $p_{low}(S)$ is the probability that the sector $S$ is underloaded
-(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_{aircraft}$:
+The simplest approch considers only the number of aircraft per sector $N_\text{aircraft}$. We
+defined two arbitary thresholds, namely $th_\text{low}$ and $th_\text{high}$, defining
+three zones where the workload is low (resp. normal and high) if $N_\text{aircraft} < th_\text{low}$
+(resp. $th_\text{low} < N_\text{aircraft} < th_\text{high}$ and $th_\text{high} < N_\text{aircraft}$).
+Those workload levels are translated to probabilities $p_\text{low}$, $p_\text{normal}$ and
+$p_\text{high}$, where $p_\text{low}(S)$ is the probability that the sector $S$ is underloaded
+(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_\text{aircraft}$:
 \begin{equation}
   \begin{aligned}
-    p_{low}^{S, t} &= \mathbb{1}_{[0, th_{low}]}(N_{aircraft}) \\
-    p_{normal}^{S, t} &= \mathbb{1}_{[th_{low}, th_{high}]}(N_{aircraft})\\
-    p_{high}^{S, t} &= \mathbb{1}_{[th_{low}, \infty[}(N_{aircraft})
+    p_\text{low}^{S, t} &= \mathbb{1}_{[0, th_\text{low}]}(N_\text{aircraft}) \\
+    p_\text{normal}^{S, t} &= \mathbb{1}_{[th_\text{low}, th_\text{high}]}(N_\text{aircraft})\\
+    p_\text{high}^{S, t} &= \mathbb{1}_{[th_\text{low}, \infty[}(N_\text{aircraft})
   \end{aligned}
@@
 use a more complex model for the workload such as a neural network
-(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_{low}$,
-$p_{normal}$ and $p_{high}$ based on parameters including trafic complexity and
+(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_\text{low}$,
+$p_\text{normal}$ and $p_\text{high}$ based on parameters including trafic complexity and
 the complexity of the sector (e.g. airspace volume).
@@
   \begin{aligned}
-    c_+(P, t) &= \sum_{S \in P_t} \delta_{h}(S, t) \cdot p_{high}^{S, t} \cdot |S|^2\\
+    c_+(P, t) &= \sum_{S \in P_t} \delta_{h}(S, t) \cdot p_\text{high}^{S, t} \cdot |S|^2\\
     c_=(P, t) &= \left(
@@
     \right)^{-1}\\
-    c_-(P, t) &= \sum_{S \in P_t} \delta_{l}(S, t) \cdot p_{low}^{S, t} \cdot |S|^{-2}
+    c_-(P, t) &= \sum_{S \in P_t} \delta_{l}(S, t) \cdot p_\text{low}^{S, t} \cdot |S|^{-2}
   \end{aligned}
@@
 with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equals 1
-if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
+if the probability $p_\text{high}$ (resp. $p_\text{normal}$ and $p_\text{low}$) is superior to
 the two others, and 0 otherwise.

### 2018-05-03 script/smooth.py
(+) smooth all the data from a given file

@@ -1,0 +1,21 @@
+import sys
+
+filename = sys.argv[1]
+
+x = []
+y = []
+
+def mean(l):
+    sum = 0
+    for i in l:
+        sum += i
+    return sum / len(l)
+
+with open(filename, 'r') as f:
+    for line in f:
+        l = line.split()
+        x.append(float(l[0]))
+        y.append(mean(list(map(float, l[1:]))))
+
+for i, j in zip(x, y):
+    print(i, ' ', j)

### 2018-05-03 script/smooth.py
(+) now in parameter the optimal value

@@ -1,6 +1,8 @@
 import sys
+import matplotlib.pyplot as plt
 
 filename = sys.argv[1]
+opt = float(sys.argv[2])
 
 x = []
 y = []
@@ -19,3 +21,7 @@ with open(filename, 'r') as f:
 
 for i, j in zip(x, y):
     print(i, ' ', j)
+
+plt.plot(x, y)
+plt.plot([min(x), max(x)], [opt, opt])
+plt.show()

### 2018-05-03 script/smooth.py
(+) with a percent line

@@ -3,6 +3,7 @@ import matplotlib.pyplot as plt
 
 filename = sys.argv[1]
 opt = float(sys.argv[2])
+percent = float(sys.argv[3]) / 100
 
 x = []
 y = []
@@ -12,6 +13,9 @@ def mean(l):
     for i in l:
         sum += i
     return sum / len(l)
+
+def fivep(x):
+    return (1+percent) * x
 
 with open(filename, 'r') as f:
     for line in f:
@@ -24,4 +28,5 @@ for i, j in zip(x, y):
 
 plt.plot(x, y)
 plt.plot([min(x), max(x)], [opt, opt])
+plt.plot([min(x), max(x)], [fivep(opt), fivep(opt)])
 plt.show()

### 2018-05-15 article/article.tex
(~) correktur
(+) Paragraph of shame

@@
 \usepackage{titlesec}
+\usepackage{tikz,pgfplots}
 
@@
   \begin{algorithmic}
-    \Function{Expand}{$v$}
+    \Function{Expand}{$u$}
     \State{}$\ell \gets \{ v | v \text{\,children of\,} u, T_v = 0 \}$
@@
 
-\subsection{Theoretical performances}
+\subsection{Theoretical performances}\label{ssec:theoretical_performances}
 In this paper, a high branching factor problem is considered.
@@
 
-TODO
+\subsection{Experimental setup}
+For the tests, the Paris West area is used. It contains 12 elementary modules
+and 20 control sectors. The resulting model tree has a mean branching factor of
+16. Since it is the smallest French area, considering the already important
+branching factor, a stochastic tree search method seems appropriate.
+
+\subsection{Accuracy}
+Accuracy is assessed by comparing the exact cost given by the A\(^*\) algorithm
+with the result of the MCTS\@. The time allowed to the MCTS to run is considered
+long enough to allow the MCTS to converge to its best solution.
+
+It results in figure~\ref{fig:longterm} that the algorithm described here has a
+bias of approximately 15\%, considering that the stabilisation matches the
+optimal value reachable.
+
+Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
+eventually converge to the optimal value, and this result reveals a flaw in the
+implementation of the algorithm.
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{axis}
+      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
+      \addplot+ table {data/longterm.data};
+      \addplot+[dashed, domain=0:10] {1268};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{The continuous line shows the total cost of a path computed by the
+    MCTS algorithm. The x-axis shows the time allowed to the algorithm. The
+    dashed curve is 15\% of the optimal cost found with the \(A^*\)
+  algorithm.}\label{fig:longterm}
+\end{figure}
+
+\subsection{Speed}
+To measure
+
+\subsection{Further work}
 

### 2018-05-15 article/article.tex
(+) More tests and further work

@@
 approach seems to give their own representation of the workload. For instance,
-~\cite{bedouet2016towards} determined workload density proportionally to the time
+\cite{bedouet2016towards} determined workload density proportionally to the time
 spent by aircraft in each sector. A simpler version~\cite{sergeeva2017dynamic}
@@
 Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
-eventually converge to the optimal value, and this result reveals a flaw in the
-implementation of the algorithm.
+eventually converge to the optimal value. This result reveals a flaw in the
+implementation.
 \begin{figure}
@@
 
-\subsection{Speed}
-To measure
+\subsection{Converging speed}
+Since the MCTS converges to a solution, one expects to know the time required to
+be close enough to the optimal solution. To have an idea of this lapse, several
+MCTS are run with different computing time. This results in a graph mapping the
+time allowed to compute each step of the MCTS to the cost of the final path. To
+get rid of the undesired varations due to the stochastic nature of the
+algorithm, the latter process is run several times; which allows one to compute
+the mean over all computed costs.
+
+The resulting graphs are shown in figure~\ref{fig:longterm} in which it can be
+seen that the algorithm reaches its best cost when it runs for 2 seconds per
+step.
+
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{axis}
+      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
+      \addplot+[dashed, domain=100:800] {18};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{The nice graph}\label{fig:timedepth}
+\end{figure}
+
+The converging speed may also depend on the depth of the tree searched through.
+The figure~\ref{fig:timedepth} shows
 
 \subsection{Further work}
+Since all the tests has been run using Paris West area, the influence of the
+complexity of the airspace has not been assessed. It would be thus relevant to
+carry further tests on bigger areas. In~\cite{ferrari2017}, the time needed to
+compute the cost with A\(^*\) depends heavily on the traffic itself. Therefore
+the use, in this paper, of randomly generated scenarii of traffic may not be
+suitable in the context of ATM. One can even use a more realistic workload model
+such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
 

### 2018-05-15 article/article.tex
(+) Scenario
(+) Abstracted

@@
 \begin{abstract}
-  A nice abstract
+  Dynamic airspace configuration falls under highly combinatorial partitioning
+  problems. Since exact methods tend to be overwhelmed by the complexity of such
+  problems, a stochastic approach is here considered. This paper presents the
+  the Monte Carlo tree search algorithm using UCT search and adapted to one
+  player games. The application of the algorithm to the airspace partitioning is
+  then detailed. Finally the Monte Carlo tree search is compared to the A\(^*\).
 \end{abstract}
@@
 
-Currently, configuration is mainly decided on the fly by chief officer (TODO).
+Currently, configuration is mainly decided on the fly by chief officer.
 This decision is based on the actual workload on each position. In this
@@
 
-TODO : meilleure explication des transitions?
-
 Each transition is composed of at most one merge, split or tranfer; more would
@@
 
+The traffic is simulated the following way, each elementary module has an
+associted list which, at index \(i\) has the number of aircraft in it at time
+step \(i\). Thoses lists has been randomly generated. The set of lists form a
+scenario which is used in all following tests.
+
 \subsection{Accuracy}

### 2018-05-18 article/article.tex
(+) Added other figure
(~) Small corrections

@@
 \begin{abstract}
-  Dynamic airspace configuration falls under highly combinatorial partitioning
-  problems. Since exact methods tend to be overwhelmed by the complexity of such
+  Dynamic airspace configuration is a highly combinatorial partitioning
+  problem. Since exact methods tend to be overwhelmed by the complexity of such
   problems, a stochastic approach is here considered. This paper presents the
-  the Monte Carlo tree search algorithm using UCT search and adapted to one
-  player games. The application of the algorithm to the airspace partitioning is
+  Monte Carlo tree search algorithm using UCT search and adapted to one
+  player games. The application of the algorithm to the airspace partitioning
+  problem is
   then detailed. Finally the Monte Carlo tree search is compared to the A\(^*\).
@@
   \item $m$: elementary module;
-  \item $S$: sector control position, group of modules;
+  \item $S$: air traffic control sector, group of modules;
   \item $P$: partition of the airspace, group of sectors;
@@
 
-Currently, configuration is mainly decided on the fly by chief officer.
+Currently, configuration is mainly decided on the fly by the chief officer.
 This decision is based on the actual workload on each position. In this
@@
 it therefore lacks a workload prediction tool.
-This papers aims at providing this tool as well as a method offering a smoother
-workload during the day.
+This papers aims at providing a method predicting the airspace configuration.
 
@@
 
-In this paper, a temporal sequence of configuration is provided, as done
+In this paper, a temporal sequence of configurations is considered, as
 in~\cite{treimuth2016branch} or~\cite{sergeeva2017dynamic}.
@@
 associated to each configuration and a transition cost. The former is based on
-the workload estimated for one sector given a neural network
+the workload estimated for one sector given by a simple model. More complex
+models might be used such as a neural network
 (see~\cite{gianazza2010forecasting}).
 
-The sequence of configurations is represented as a tree. The resulting
+The sequence of configurations is extracted from a tree. The resulting
 sequence will therefore be a path from the root to a leaf of this tree. This
@@
 
-The dynamic airspace configuration problem requires a model of the airspace. As
-done in~\cite{sergeeva2017dynamic} or~\cite{treimuth2016branch} the airspace
+The dynamic airspace configuration problem requires a model of the airspace.
+In~\cite{sergeeva2017dynamic} or~\cite{treimuth2016branch} the airspace
 is modelled via a graph. In those graphs, vertices represent elementary modules
@@
 This way any configuration will match specified requirements, which can be
-qualified hard constraints.
+qualified as hard constraints.
 
@@
 volume, or vertical incoming flows in the next 15 and 60 minutes. Those
-indicators ease the training of a neural network forecasting the workload.
+indicators as inputs to a neural network forecasting the workload.
 
@@
 configurations, the work associated with the reallocation of one or more modules
-is evaluated. This can be used in a cost function aimed to be minimised
-(in~\cite{bedouet2016towards}).
+is evaluated. This transition cost is included in the cost function being
+minimised (in~\cite{bedouet2016towards}).
 
@@
 
-The Monte Carlo tree search algorithm has been quickly used in two players game.
+The Monte Carlo tree search algorithm has been widely used in two-player games.
 Only three years after its apparition in 1990, Br\"ugmann applies it to the Go
 game in~\cite{brugmann1993mcgo}. While the Monte-Carlo tree search is still
-extensively used in two players games (and especially the Go game), its
+extensively used in two-player games (and especially the Go game), its
 adaptation to one player game has been worked on. Auer \textit{et al.} propose
 an
-upper confidence bound formula which appears to be more efficient for one player
-games as seen in~\cite{sebag2010fuse}. The introduced formula uses the standard
+upper confidence bound formula in~\cite{sebag2010fuse} which appears to be more
+efficient for one player games. The introduced formula uses the standard
 deviation of the outcome of the simulations.
@@
 stochastic tree search method, namely the Monte Carlo tree search. To assess the
-exactness of the results, an exact method (here A\(^*\)) is used.
+quality of the results, an exact method (here A\(^*\)) is used.
 
@@
 the previously mentioned policies. Then the MCTS algorithm is called back with
-node \(u\)
-as the root node. The path is composed of the successive roots. The algorithm is
-described at~\ref{alg:path_building}.
+node \(v\)
+as the root node. The final path is composed of the successive roots. The
+algorithm is described in~\ref{alg:path_building}.
 \begin{algorithm}
@@
 To evaluate the exactness of the paths given by the MCTS algoithm, the A\(^*\)
-algorithm is used. A\(^*\) is an exact depth first search pathfinding algorithm
+algorithm is used. A\(^*\) is an exact best first search pathfinding algorithm
 which uses a heuristic function to guide its search. The algorithm is given
-in~\ref{alg:astar} where \(u_0\) is the initial state, \(T\) the set of termial
+in~\ref{alg:astar} where \(u_0\) is the initial state, \(T\) the set of terminal
 nodes, \(h\) a heuristic function estimating the cost from a state \(u\) given
@@
 In this section, we discuss the model established to approach this configuration
-problem. First we defined a structure for a controled sector and all possible
+problem. First we define a structure for a controlled sector and all possible
 transitions through time. Then a workload model is defined. And finally we
-determined a cost per partition and a cost function aimed to be minimized.
+define a cost per partition and a cost function aimed to be minimized.
 
@@
 one or more elementary modules. A partition $P$ is a set of
-sectors covering all the airspace.
+non-overlapping sectors covering all the airspace.
 
 A sector is a group of elementary modules, but in practice only a collection
-of sectors are allowed in this model. This collection is referred as the context.
-This restriction reflects the fact the more complex the shape is, the harder it
-is to manage.
+of sectors are allowed in this model. This collection is referred to as the
+context. It is the list of operation ATC sectors actually used in ATC centres.
 
@@
 
-During the day, and with the evolution of the trafic, the workload (rigorously
-defined in~\ref{sub:workload}) vary. To help controllers to maintain a
+During the day, and with the evolution of the traffic, the workload (rigorously
+defined in~\ref{sub:workload}) varies. To help controllers to maintain a
 homogeneous level of workload, the airspace partitioning needs to be adapted
-through a transition from an inital partition to a target partition.
+through successive transitions starting from an inital partition.
 This maneuver requires either coordination or the opening of a new position
@@
 
-In the model considered here, three transistions are possible namely a merge,
+In the model considered here, three transitions are possible namely a merge,
 a transfer
@@
 
-Each transition is composed of at most one merge, split or tranfer; more would
+Each transition is composed of at most one merge, split or transfer; more would
 result in too complex reconfigurations.
@@
 Let us now precise the model established to evaluate a workload metric. The
-workload depends on external criteria such as sector and trafic complexity
-or humain factors (e.g. controller health or stress). In the model assumed here,
-the workload felt by a controller is directly and only related to the trafic
+workload depends on external criteria such as sector and traffic complexity
+or human factors (e.g.\ controller health or stress). In the model assumed here,
+the workload felt by a controller is directly and only related to the traffic
 complexity.
@@
 The simplest approch considers only the number of aircraft per sector $N_\text{aircraft}$. We
-defined two arbitary thresholds, namely $th_\text{low}$ and $th_\text{high}$, defining
+define two arbitary thresholds, namely $th_\text{low}$ and $th_\text{high}$, defining
 three zones where the workload is low (resp. normal and high) if $N_\text{aircraft} < th_\text{low}$
-(resp. $th_\text{low} < N_\text{aircraft} < th_\text{high}$ and $th_\text{high} < N_\text{aircraft}$).
+(resp. $th_\text{low} \leq N_\text{aircraft} \leq th_\text{high}$ and $th_\text{high} < N_\text{aircraft}$).
 Those workload levels are translated to probabilities $p_\text{low}$, $p_\text{normal}$ and
@@
 
-This expression of the workload, expressed in term of probabilities, allow us to
+This expression of the workload, expressed in terms of probabilities, allows us to
 use a more complex model for the workload such as a neural network
 (see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_\text{low}$,
-$p_\text{normal}$ and $p_\text{high}$ based on parameters including trafic complexity and
+$p_\text{normal}$ and $p_\text{high}$ based on parameters including traffic complexity and
 the complexity of the sector (e.g. airspace volume).
@@
 In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
-to be defined. In this paper, this cost depends on the workload in each the
+to be defined. In this paper, this cost depends on the workload in each
 sector. The definitions given in~\cite{ferrari2017} are used to represent a high
@@
 value $\alpha$ help to
-minimize $c_+$, hence the overall number of sectors with too much trafic.
+minimize $c_+$, hence the overall number of sectors with too much traffic.
 In a real application, it maybe interesting to order those parameters as
@@
 with $\theta > 0$ a parameter to determine.
-This is the immediate function to minimize in the A$^*$ algorithm.
+This is the objective function to minimize in the A$^*$ algorithm.
 
@@
 associted list which, at index \(i\) has the number of aircraft in it at time
-step \(i\). Thoses lists has been randomly generated. The set of lists form a
+step \(i\). Those lists have been randomly generated. The set of lists form a
 scenario which is used in all following tests.
@@
   algorithm.}\label{fig:longterm}
+\end{figure}
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{axis}
+      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
+      \addplot+ table {data/shortterm.data};
+      \addplot+[dashed, domain=0:2] {1268};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{}\label{fig:shortterm}
 \end{figure}

### 2018-05-18 article/article.tex
(+) new figure with caption

@@
 An other way to model the problem is to use a constrained set of
-configurations as in~\cite{gianazza2010forecasting} and~\cite{bloem2010dynamic}.
+configurations as in~\cite{bloem2010dynamic}.
 This way any configuration will match specified requirements, which can be
@@
 each
-elementary module and the cost of transition from the previous partitioning. The
-available partitions come from a predetermined set containing feasible
-partitioning.
+elementary module and the cost of transition from the previous partitioning.
+The set of available parititions can be computed from the set of elementary
+modules and a context which contains all available ATC sectors.
 
@@
 Since the overall goal is to compute a temporal sequence of partitions over a
-day, the day has to be sampled. This sampling in represented as the temporal
-difference between a node and its children in the trees mentioned in the
-previous section.
+day, the day has to be sampled. In the model tree, a timestep separates a node
+from its children. Practically, a timestep is one minute.
 
@@
 To use effectively the A\(^*\) algorithm, a heuristic is needed. Using the null
-heuristic (\(\forall u \in \mathcal{N}, h(u) = 0\)) revealed to be inefficient
-considering the branching factor. To build a minimal heuristic, the limitations
-introduced in~\ref{sssec:transitions} are not considered, neither those associated
-to the previous partition (split, merge or transfer). This allows to generate
+heuristic (\(\forall u \in \mathcal{N}, h(u) = 0\)) -- and thus using Dijkstra
+algorithm -- revealed to be inefficient
+considering the branching factor. Let \(k \in \mathbb{N}\), \(u\) a node
+selected at time \(k\).
+The heuristic \(h(u)\) gives the cost of the path starting from \(u\) to a
+leaf which is reached at a time \(t_f\).
+The heuristic is built without
+considering transitions (transition cost and~\ref{sssec:transitions}).
+Let \(P^*_i\) be the partition with minimal cost at time \(i\), the heuristic is
+then
+\begin{equation}
+  h(u) = \sum_{i=k+1}^f C(P^*_i, t_i)
+\end{equation}
+This allows to generate
 the sequence of the best partitions among all possible at each time step.
@@
       \addplot+ table {data/shortterm.data};
-      \addplot+[dashed, domain=0:2] {1268};
     \end{axis}
   \end{tikzpicture}
-  \caption{}\label{fig:shortterm}
+  \caption{Monte Carlo tree search called with varying time allowed to compute
+  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
+  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
 \end{figure}

### 2018-05-18 article/article.tex
(~) Reworked results

@@
 \usepackage{tikz,pgfplots}
+\usepackage{array,tabulary,booktabs}
 
@@
 \subsection{Experimental setup}
-For the tests, the Paris West area is used. It contains 12 elementary modules
-and 20 control sectors. The resulting model tree has a mean branching factor of
-16. Since it is the smallest French area, considering the already important
+For the tests, a fictitious area is used. It contains 5 elementary modules
+and 12 control sectors. The resulting model tree has a mean branching factor of
+6.
+Regarding the branching, the smallest French area -- which is Paris West
+-- has a branching factor of 16 for 20 control sectors and 12 elementary
+modules. Considering the already important
 branching factor, a stochastic tree search method seems appropriate.
@@
 associted list which, at index \(i\) has the number of aircraft in it at time
-step \(i\). Those lists have been randomly generated. The set of lists form a
-scenario which is used in all following tests.
+step \(i\). Those lists have been written by hand, the activity shows a traffic
+peak at the beginning of the scenario and much less activity after. The expected
+outcome of the algorithms is a massive degrouping reflecting the heavy workload
+followed by merges.
 
 \subsection{Accuracy}
-Accuracy is assessed by comparing the exact cost given by the A\(^*\) algorithm
-with the result of the MCTS\@. The time allowed to the MCTS to run is considered
-long enough to allow the MCTS to converge to its best solution.
+Accuracy is assessed by comparing the outcome of the MCTS algorithm with the
+A\(^*\) cost and a greedy algorithm cost.
+The time allowed to the MCTS to run is considered long enough to allow the MCTS
+to converge to its best solution. The results are summarised in
+table~\ref{tab:methods_compare}.
 
-It results in figure~\ref{fig:longterm} that the algorithm described here has a
-bias of approximately 15\%, considering that the stabilisation matches the
-optimal value reachable.
+The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
+A\(^*\) and MCTS regarding grouping and degrouping.
 
@@
 implementation.
+
+\begin{table}
+  \centering
+  \begin{tabulary}
+    {\textwidth}{LRR}
+    Alg. & Cost & Err.\\\toprule
+    A\(^*\) & 77.31 & 0\%    \\
+    MCTS    & 82.48 & 6.7\%  \\
+    Greedy  & 90.50 & 17.1\% \\
+  \end{tabulary}
+  \caption{Comparison of several methods on the generated
+  scenario.}\label{tab:methods_compare}
+\end{table}
+
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{axis}
+      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
+      legend entries={A\(^*\), MCTS}]
+      \foreach \i in {0,1}{%
+        \addplot+ table [
+          x expr={\lineno}, y index=\i
+          ]
+          {./data/grouping.data};
+      }
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Grouping profile}\label{fig:grouping}
+\end{figure}
 \begin{figure}

### 2018-05-19 article/article.tex
(+) New par

@@
 
+\subsubsection*{Tree translation}
+In the tree used in the algorithm exposed earlier, each node has an embedded
+state. Considering a node \(u\), its children are the elements of the set of
+nodes having as states the possible transitions from the state of \(u\). Going
+down in the tree by selecting nodes is then equivalent to moving forward through
+time, each step in the tree being a time step.
+
 \subsection{Workload}\label{sub:workload}
@@
 $p_\text{normal}$ and $p_\text{high}$ based on parameters including traffic complexity and
-the complexity of the sector (e.g. airspace volume).
+the complexity of the sector (e.g.\ airspace volume).
 
@@
 the use, in this paper, of randomly generated scenarii of traffic may not be
-suitable in the context of ATM. One can even use a more realistic workload model
-such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
+suitable in the context of ATM\@. One can even use a more realistic workload
+model such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
 

### 2018-05-19 article/article.tex
(~) Heading

@@
 \renewcommand{\thesection}{\Roman{section}}
-\renewcommand{\thesubsection}{\Alph{subsection}}
+\renewcommand{\thesubsection}{\Roman{section}.\Alph{subsection}}
 
@@
 define two arbitary thresholds, namely $th_\text{low}$ and $th_\text{high}$, defining
-three zones where the workload is low (resp. normal and high) if $N_\text{aircraft} < th_\text{low}$
+three zones where the workload is low (resp.\ normal and high) if $N_\text{aircraft} < th_\text{low}$
 (resp. $th_\text{low} \leq N_\text{aircraft} \leq th_\text{high}$ and $th_\text{high} < N_\text{aircraft}$).
@@
 $p_\text{high}$, where $p_\text{low}(S)$ is the probability that the sector $S$ is underloaded
-(resp. balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_\text{aircraft}$:
+(resp.\ balanced and overloaded). In short, for a sector $S$ at time $t$ and $N_\text{aircraft}$:
 \begin{equation}

### 2018-05-19 article/article.tex
(+) Corr in article and continued slideshow

@@
   \begin{aligned}
-    p_\text{low}^{S, t} &= \mathbb{1}_{[0, th_\text{low}]}(N_\text{aircraft}) \\
+    p_\text{low}^{S, t} &= \mathbb{1}_{[0, th_\text{low}[}(N_\text{aircraft}) \\
     p_\text{normal}^{S, t} &= \mathbb{1}_{[th_\text{low}, th_\text{high}]}(N_\text{aircraft})\\
-    p_\text{high}^{S, t} &= \mathbb{1}_{[th_\text{low}, \infty[}(N_\text{aircraft})
+    p_\text{high}^{S, t} &= \mathbb{1}_{]th_\text{low}, \infty[}(N_\text{aircraft})
   \end{aligned}

### 2018-05-20 article/article.tex
(+) More detail on degrouping
(+) Added data

--- 2018-05-19 article/article.tex
@@
 
-\subsubsection*{Tree translation}
-In the tree used in the algorithm exposed earlier, each node has an embedded
-state. Considering a node \(u\), its children are the elements of the set of
-nodes having as states the possible transitions from the state of \(u\). Going
-down in the tree by selecting nodes is then equivalent to moving forward through
-time, each step in the tree being a time step.
-
 \subsection{Workload}\label{sub:workload}
@@
 $p_\text{normal}$ and $p_\text{high}$ based on parameters including traffic complexity and
-the complexity of the sector (e.g.\ airspace volume).
+the complexity of the sector (e.g. airspace volume).
 
@@
 
-Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
-eventually converge to the optimal value. This result reveals a flaw in the
-implementation.
-
 \begin{table}
@@
       [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
-      \addplot+ table {data/longterm.data};
-      \addplot+[dashed, domain=0:10] {1268};
-    \end{axis}
-  \end{tikzpicture}
-  \caption{The continuous line shows the total cost of a path computed by the
-    MCTS algorithm. The x-axis shows the time allowed to the algorithm. The
-    dashed curve is 15\% of the optimal cost found with the \(A^*\)
-  algorithm.}\label{fig:longterm}
-\end{figure}
-\begin{figure}
-  \begin{tikzpicture}
-    \begin{axis}
-      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
       \addplot+ table {data/shortterm.data};
@@
 
+The converging speed may also depend on the depth of the tree searched through.
+The figure~\ref{fig:timedepth} shows
+
 \begin{figure}
@@
       [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
-      \addplot+[dashed, domain=100:800] {18};
+      \foreach \i in {1, 2} {%
+        \addplot+ table [ x index = 0, y index = \i ] {./data/cmpdepth.data};
+      }
     \end{axis}
   \end{tikzpicture}
-  \caption{The nice graph}\label{fig:timedepth}
+  \caption{Cost difference for varying depth with 0.5 seconds allowed to MCTS,
+  \(\theta=3\)}\caption{fig:cmpdepth}
 \end{figure}
-
-The converging speed may also depend on the depth of the tree searched through.
-The figure~\ref{fig:timedepth} shows
 
@@
 the use, in this paper, of randomly generated scenarii of traffic may not be
-suitable in the context of ATM\@. One can even use a more realistic workload
-model such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
+suitable in the context of ATM. One can even use a more realistic workload model
+such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
 

### 2018-05-20 article/article.tex
(+) Added data

--- 2018-05-19 article/article.tex
@@
 The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
-A\(^*\) and MCTS regarding grouping and degrouping.
-
-Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
-eventually converge to the optimal value. This result reveals a flaw in the
-implementation.
+A\(^*\) and MCTS regarding grouping and degrouping. On the \(y\) axis is
+indicated the number of sectors. The two algorithms show, as expected, splits
+leading from 3 sectors to 5 for the MCTS to cope with the intense trafic at
+the beginning of the scenario. Both algorithm choose to merge as much as
+possible to go to 1 sector.
 

### 2018-05-20 article/article.tex
(+) Nice conclusion

@@
 
-\subsection{Further work}
-Since all the tests has been run using Paris West area, the influence of the
-complexity of the airspace has not been assessed. It would be thus relevant to
-carry further tests on bigger areas. In~\cite{ferrari2017}, the time needed to
-compute the cost with A\(^*\) depends heavily on the traffic itself. Therefore
-the use, in this paper, of randomly generated scenarii of traffic may not be
-suitable in the context of ATM\@. One can even use a more realistic workload
-model such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
+\section{Conclusion}
+The presented work provides the basis of a method able to create dynamic
+airspace overture schemes. The model is built around a quantification of the
+workload felt by a air trafic controller. This quantification is then used to
+create a cost function to be minimised. The highly combinatorial nature of the
+problem makes it prone to be solved with stochastic methods. The Monte Carlo
+tree search algorithm is then introduced since it has been used in some of the
+most combinatorial problem (game of Go). In this paper, the algorithm is
+modified to take into account not only a result but all the path leading to this
+leaf. It is also adapted to the one-player game nature of the problem.
+
+\paragraph{Further work}
+Since all the tests has been run using the same fictitious area with the same
+trafic, neither the influence of the complexity of the airspace nor of the
+trafic have been assessed. It would be thus relevant to carry further tests on
+bigger areas with differing trafic. This can be emphasised
+by~\cite{ferrari2017}, in which the time needed to compute the cost with A\(^*\)
+depends heavily on the trafic itself.
+
+More complex workload models can also be used such as neural networks
+in~\cite{gianazza2010forecasting}.
 

### 2018-05-20 article/article.tex
(+) Par on grouping results

@@
 associted list which, at index \(i\) has the number of aircraft in it at time
-step \(i\). Those lists have been written by hand, the activity shows a traffic
-peak at the beginning of the scenario and much less activity after. The expected
-outcome of the algorithms is a massive degrouping reflecting the heavy workload
-followed by merges.
+step \(i\). Those lists have been written by hand.
 
@@
 A\(^*\) and MCTS regarding grouping and degrouping. On the \(y\) axis is
-indicated the number of sectors. The two algorithms show, as expected, splits
-leading from 3 sectors to 5 for the MCTS to cope with the intense trafic at
-the beginning of the scenario. Both algorithm choose to merge as much as
-possible to go to 1 sector.
-
+indicated the number of sectors. The trafic activity is the following,
+\begin{enumerate}
+  \item\label{scen:busy}
+    The first part of the scenario shows a busy situation, all sectors are
+    overloaded;
+  \item\label{scen:quiet}
+    then follows a quiet moment with few aircraft, much of the sectors are
+    underloaded;
+  \item\label{scen:partial}
+    the end of the scenario has two overload modules and the rest under
+    normal or under load.
+\end{enumerate}
+The expected behaviour is a tendency to increase the number of sectors through
+splits to cope with part~\ref{scen:busy} followed by many merges to reduce the
+number of sectors in response to phase~\ref{scen:quiet}. The algorithms should
+end with a specific sector created, the one overloaded in~\ref{scen:partial} and
+the rest grouped. Seeing~\ref{fig:grouping}, both algorithm behave as expected.
 \begin{table}

### 2018-05-20 article/article.tex
(+) Introduced new results

@@
 \subsection{Accuracy}
-Accuracy is assessed by comparing the outcome of the MCTS algorithm with the
-A\(^*\) cost and a greedy algorithm cost.
-The time allowed to the MCTS to run is considered long enough to allow the MCTS
-to converge to its best solution. The results are summarised in
-table~\ref{tab:methods_compare}.
-
 The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
@@
 the rest grouped. Seeing~\ref{fig:grouping}, both algorithm behave as expected.
-\begin{table}
-  \centering
-  \begin{tabulary}
-    {\textwidth}{LRR}
-    Alg. & Cost & Err.\\\toprule
-    A\(^*\) & 77.31 & 0\%    \\
-    MCTS    & 82.48 & 6.7\%  \\
-    Greedy  & 90.50 & 17.1\% \\
-  \end{tabulary}
-  \caption{Comparison of several methods on the generated
-  scenario.}\label{tab:methods_compare}
-\end{table}
-
 \begin{figure}
@@
 \end{figure}
+
+Accuracy can be quantitatively assessed by comparing the outcome of the MCTS
+algorithm against the A\(^*\) cost and a greedy algorithm cost.
+The time allowed to the MCTS to run is considered long enough to allow the MCTS
+to converge to its best solution. The results are summarised in
+table~\ref{tab:methods_compare}.
+\begin{table}
+  \centering
+  \begin{tabulary}
+    {\textwidth}{LRR}
+    Alg.    & Cost  & Err.   \\
+    \toprule
+    A\(^*\) & 77.31 & 0\%    \\
+    MCTS    & 82.48 & 6.7\%  \\
+    Greedy  & 90.50 & 17.1\% \\
+  \end{tabulary}
+  \caption{Comparison of several methods on the generated
+  scenario.}\label{tab:methods_compare}
+\end{table}
+
+The graph~\ref{fig:cmpdepth} shows the evolution of the difference of costs with
+varying depth. The error rate is expected to grow as the depth increase since
+the MCTS might fail to explore the best nodes among the exponentially growing
+amount of them.
 \begin{figure}
@@
     \begin{axis}
-      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
-      \addplot+ table {data/shortterm.data};
+      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost,
+      legend entries = {A\(^*\), MCTS}]
+      \foreach \i in {0, 1} {%
+        \addplot+ table [
+          x expr = {\lineno + 1}, y index = \i ] {./data/cmpdepth.data};
+      }
     \end{axis}
   \end{tikzpicture}
-  \caption{Monte Carlo tree search called with varying time allowed to compute
-  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
-  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
+  \caption{Cost difference for varying depth with 0.5 seconds allowed to
+    MCTS.}\label{fig:cmpdepth}
 \end{figure}
@@
 The figure~\ref{fig:timedepth} shows
-
 \begin{figure}
@@
     \begin{axis}
-      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
-      \foreach \i in {1, 2} {%
-        \addplot+ table [ x index = 0, y index = \i ] {./data/cmpdepth.data};
-      }
+      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
+      \addplot+ table {data/shortterm.data};
     \end{axis}
   \end{tikzpicture}
-  \caption{Cost difference for varying depth with 0.5 seconds allowed to MCTS,
-  \(\theta=3\)}\caption{fig:cmpdepth}
+  \caption{Monte Carlo tree search called with varying time allowed to compute
+  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
+  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
 \end{figure}

### 2018-05-21 article/article.tex
(~) Small corrections

@@
 the MCTS might fail to explore the best nodes among the exponentially growing
-amount of them.
+amount of them. The figure~\ref{fig:cmpdepth} does not show a constant growth of
+error rate between MCTS and A\(^*\). As hoped, the MCTS remains always better
+than the greedy algorithm.
 \begin{figure}
@@
       [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost,
-      legend entries = {A\(^*\), MCTS}]
-      \foreach \i in {0, 1} {%
+      legend entries = {A\(^*\), MCTS, Greedy}, legend pos=south east]
+      \foreach \i in {0, 1, 2} {%
         \addplot+ table [
@@
 
-The resulting graphs are shown in figure~\ref{fig:longterm} in which it can be
-seen that the algorithm reaches its best cost when it runs for 2 seconds per
-step.
-
-The converging speed may also depend on the depth of the tree searched through.
-The figure~\ref{fig:timedepth} shows
+The resulting graphs are shown in figure~\ref{fig:shortterm} in which it can be
+seen that the algorithm reaches its best cost when it runs for 1 second per
+step, but the cost is approximately stable from 0.5 seconds.
 \begin{figure}
@@
 trafic have been assessed. It would be thus relevant to carry further tests on
-bigger areas with differing trafic. This can be emphasised
+bigger areas with heavier and differing trafic. This can be emphasised
 by~\cite{ferrari2017}, in which the time needed to compute the cost with A\(^*\)

### 2018-05-22 article/article.tex

@@
   \caption{Comparison of several methods on the generated
-  scenario.}\label{tab:methods_compare}
+  scenario (\(\theta = 3\)).}\label{tab:methods_compare}
 \end{table}
@@
   \caption{Cost difference for varying depth with 0.5 seconds allowed to
-    MCTS.}\label{fig:cmpdepth}
+  MCTS and \(\theta = 3\).}\label{fig:cmpdepth}
 \end{figure}
